---
title: "TFM2_16S_amplicons_analysis"
author: "lrubinat"
date: "5/30/2017"
output:
  html_document:
    theme: united
    toc: yes
  pdf_document:
    highlight: zenburn
    toc: yes
---


<!--- INITIALIZATION
```{r, echo=FALSE}
#error hook to kill knitr in case of errors
library(knitr)
knit_hooks$set(error = function(x, options) stop(x))
opts_chunk$set(cache=TRUE, autodep=TRUE)
```
--->

```{r libraries, echo=F, results="hide", warning=F, message=F}
library(data.table)
library(dplyr)
library(DT)
library(ggplot2)
library(gtools)
library(vegan)
library(iNEXT)
library(fossil)
library(ggrepel)
library(dendextend)
library(forcats)
```


# 1) 16S amplicons

## 1.1) Data overview


``` {r load_data, echo=FALSE, message=FALSE, results="hide", warning=F}
#Let's read the dataset and remove the samples containing less than 16509 reads:

setwd("~/Documents/TFM2/TFM2_16S_amplicons_analysis/")

#read data 
tb16_tax <- read.table(file="~/Documents/TFM2/TFM2_16S_amplicons_analysis/input_99/OTUtb_16S_NR99_curated_9114_OTUs_nospchar_v2.txt", head=TRUE, fill=TRUE, sep="\t")

#table dimensions and format before setting column names
dim(tb16_tax) # 9114  131
tb16_tax[1:5,1:5]

row.names(tb16_tax)<-tb16_tax[,1]
tb16_tax<-tb16_tax[,-1]
tb16_tax[is.na(tb16_tax)]<-"NA"

dim(tb16_tax) #9114  130
tb16_tax[1:5,1:5]

#drop repeated samples in stations 68a and 71a:
tb16_tax<-subset(tb16_tax, select=-c(st068a_MD1254,st071a_MD1318))
#remove OTUs that might only occurr in samples 68a and 71a:
#tb16_tax<-tb16_tax[-(which(rowSums(tb16_tax[,1:122])==0)),]
dim(tb16_tax)

#table with taxonomic classification alone:
tb16_class <- tb16_tax[,123:128]
dim(tb16_class) #9114   7
tb16_class[1:5,1:6]

#table with occurence data alone
tb16_tax_occur <- tb16_tax[,1:122]
dim(tb16_tax_occur) #9114   122
tb16_tax_occur[1:5,1:5]
```

The dataset "Taraspina 16S miTags" contains reads from **122 samples of Malaspina**. On average, each sample contains 45940 reads:

```{r summary, echo=FALSE, message=FALSE, warning=F}
amplicons_per_sample_tb16<-colSums(tb16_tax_occur)
summary(amplicons_per_sample_tb16) 
#median of amplicons per sample = ~33100)
```

```{r samples_selection, echo=FALSE, message=FALSE, results="hide", warning=F}
#we'll remove all samples with less than 16509 reads.
sort(amplicons_per_sample_tb16[which(colSums(tb16_tax_occur)<16509)])

#remove samples with less than 16509 reads
tb16_tax_occur_min16509 <- tb16_tax_occur[,colSums(tb16_tax_occur) >= 16509]
dim(tb16_tax_occur_min16509) #9114 106

#remove samples with omitted in 18S dataset (so that we can compare the relative abundance of 18S and 16S OTUs considering the same samples)
tb16_tax_occur_min16509<-subset(tb16_tax_occur_min16509, select=-c(st056_MD1035,st063_MD1176,st064_MD1186,st067_MD1246,st083_MD1517,st013_MD202,st029_MD506,st030_MD528,st004_MD60,st037_MD646,st046_MD834,st051_MD916,st053_MD962,st054_MD985,st007_MD98))

dim(tb16_tax_occur_min16509) #9114  91

#let's check if we loose any OTU when excluding the mentioned stations.
tb16_tax_occur_min16509<-tb16_tax_occur_min16509[-(which(rowSums(tb16_tax_occur_min16509)==0)),]
dim(tb16_tax_occur_min16509)
```


```{r starting_dataset, echo=FALSE, results="hide"}
#Table dimensions and content outline:

dim(tb16_tax_occur_min16509)
#tb16_tax_occur_min16509[1:5,1:5]
```


```{r reads_per_sample_overview1, echo=F, results="hide"}
#Minimum number of reads per station:

min(colSums(tb16_tax_occur_min16509)) 
#16509
```


```{r reads_per_sample_overview2, echo=F, results="hide"}
#Maximum number of reads per station:

max(colSums(tb16_tax_occur_min16509)) 
# max: 148202
```


```{r reads_per_sample_overview3, echo=F, results="hide"}
#Identification of stations with higher number of reads:

amplicons_per_sample<-colSums(tb16_tax_occur_min16509)
amplicons_per_sample[which(colSums(tb16_tax_occur_min16509)>140000)]
```

\
Overall reads per sample:

``` {r reads_per_sample_overview4, echo=FALSE}
plot(sort(colSums(tb16_tax_occur_min16509)), pch=19, xlab="sample no.", ylab="reads per sample", cex=0.9)
```


## 1.2) Normalization

\
In order to keep as many samples as possible, we rarefy at **16509 reads per sample**. By that, we loose 31 samples, and after removing the exluded samples in the 16S dataset (to make them comparable), we end up with a **normalized dataset containing 91 samples and 8173 OTUs**.


``` {r species_richness_rarefaction1, echo=FALSE, results="hide"}
#Let's normalize the original dataset by randomly subsampling 16509 reads in each station:

tb16_tax_occur_min16509_t<-t(tb16_tax_occur_min16509)
tb16_tax_occur_ss16509<-rrarefy(tb16_tax_occur_min16509_t, 16509)
```


```{r species_richness_rarefaction2, echo=FALSE, results="hide"}
dim(tb16_tax_occur_ss16509)
#tb16_tax_occur_ss16509[1:5,1:5]
```


``` {r species_richness_rarefaction3, echo=F, results="hide"}
#Its content fits with the expected normalization values (16509 reads per station):

rowSums(tb16_tax_occur_ss16509)
```


```{r species_richness_rarefaction4, echo=F, results="hide"}
#OTUs we loose in the new table:

length(which(colSums(tb16_tax_occur_ss16509)==0)) #684
```


```{r species_richness_rarefaction5, echo=F, results="hide"}
##let's remove them from the table and take a look at its final dimensions:

tb16_tax_occur_ss16509_no_cero<-tb16_tax_occur_ss16509[,-(which(colSums(tb16_tax_occur_ss16509)==0))]
tb16_tax_occur_ss16509_no_cero<-tb16_tax_occur_ss16509_no_cero[mixedorder(row.names(tb16_tax_occur_ss16509_no_cero)),]
dim(tb16_tax_occur_ss16509_no_cero) #91 8154

#the final dimensions of the normalized table are 91 8197.
#8197 + 684 = 8881
```

\
Datasets summary:
```{r dataset_summary,  echo=T}
dim(tb16_tax) #original dataset

dim(tb16_tax_occur) #original dataset with occurrence data alone

dim(tb16_tax_occur_min16509) #dataset without samples with less than 16509 OTUs

dim(tb16_tax_occur_ss16509_no_cero) #rarefied dataset
```


## 1.3) General community analysis

### 1.3.1) Richness and evenness (Shannon index)

```{r shannon_index1, echo=FALSE}
tb16_tax_occur_ss16509_div <- diversity(tb16_tax_occur_ss16509_no_cero, index="shannon")
```

Most of the samples take Shannon Index values around  4:

```{r shannon_index2, echo=FALSE}
boxplot(tb16_tax_occur_ss16509_div, pch=19, main="MP 16S Shannon's index of diversity",ylab="Shannon's index")
plot(sort(tb16_tax_occur_ss16509_div), pch=19, main="MP 16S Shannon's index of diversity",ylab="Shannon's index")
```

### 1.3.2) Richness: OTU number

```{r richness_otu_no1, echo=FALSE}
OTUs_per_sample_16S_tax_occur_ss16509<-specnumber(tb16_tax_occur_ss16509_no_cero)
```

Lowest number of OTUs per sample:

```{r richness_otu_no2, echo=FALSE}
min(OTUs_per_sample_16S_tax_occur_ss16509)
```

Maximum number of OTUs per sample:

```{r richness_otu_no3, echo=FALSE}
max(OTUs_per_sample_16S_tax_occur_ss16509)
```

In most of the samples, we can identify about 1400 OTUs:

```{r richness_otu_no4, echo=F}
plot(sort(OTUs_per_sample_16S_tax_occur_ss16509), pch=19, main="MP 16S no. of OTUs per sample", xlab="sample no.", ylab="no. of OTUs")
boxplot(OTUs_per_sample_16S_tax_occur_ss16509, pch=19, main="MP 16S no. of OTUs per sample", ylab="no. of OTUs")
```

### 1.3.3) Index of evenness

#### 1.3.3.1) Pielou's index

```{r pielou_index_of_evenness1, echo=F, results="hide"}
pielou_evenness_16S_tax_occur_ss16509<-tb16_tax_occur_ss16509_div/log(OTUs_per_sample_16S_tax_occur_ss16509)
```

The Pielou index (constrained between 0 and 1) takes values closer to 1 as the variation of species proportion in a sample increases. Our samples get values around 0.6, meaning that the numerical composition of different OTUs in a sample is not so variable - we might observe certain dominant species.

```{r pielou_index_of_evenness2, echo=F}
plot(sort(pielou_evenness_16S_tax_occur_ss16509), pch=19, main="TP 16S Pielou's index", xlab="sample no.", ylab="Pielou's index")
boxplot(pielou_evenness_16S_tax_occur_ss16509,  pch=19, main="TP 16S Pielou's index", xlab="sample no.", ylab="Pielou's index")
```


<!---
#### 1.3.3.2) Sads
library(sads)
?sads
--->

### 1.3.4) Abundance Models


```{r OTUs_overall_abundance, echo=F, results="hide"}
#The OTU_1, with 409610 reads, is the most abundant in the overall dataset:

head(sort(colSums(tb16_tax_occur_ss16509_no_cero), decreasing=T), n=10L)
```

Most of the OTUs show very few occurrences, suggesting that we will probably be able to identify a significant ammount of rare otus:

```{r OTUs_overall_abundance2, echo=F, results="hide"}
#previous version:
#plot(log(sort(colSums(tb16_tax_occur_ss16509_no_cero), decreasing=T)), pch=19)

tb16_tax_occur_ss16509_no_cero_rank<-as.data.table(t(tb16_tax_occur_ss16509_no_cero))
tb16_tax_occur_ss16509_no_cero_rank[,V1:=rowSums(tb16_tax_occur_ss16509_no_cero_rank)]
tb16_tax_occur_ss16509_no_cero_rank<-tb16_tax_occur_ss16509_no_cero_rank[order(V1,decreasing = T)]
tb16_tax_occur_ss16509_no_cero_rank[,rank:=1:nrow(tb16_tax_occur_ss16509_no_cero_rank)]

p1<-ggplot(tb16_tax_occur_ss16509_no_cero_rank, aes(x=rank,y=V1))+
  geom_point(aes(x=rank,y=V1))+
  scale_y_log10(name="abundance (log scale)")+
  labs(title="TP 16S rank abundance curve")+
  theme_bw()+
  labs(colour = "") 
p1
```



#### 1.3.4.1) Rank-Abundance or Dominance/Diversity Model ("radfit")

The OTUs abundance distribution fits relativelly close to log-normal model. 

```{r radfit, echo=FALSE}
#?radfit
#otu_tb16_t[1:5,1:5]

tb16_tax_occur_min16509_radfit<-radfit(colSums(tb16_tax_occur_min16509_t))
plot(tb16_tax_occur_min16509_radfit)
```

#### 1.3.4.2) Preston's Lognormal Model

According to Preston's lognormal model fit into species frequencies groups, we're missing ~262 species:

```{r preston_model1, echo=4}
tb16_tax_occur_ss16509_prestonfit<-prestonfit(colSums(tb16_tax_occur_min16509_t))
plot(tb16_tax_occur_ss16509_prestonfit, main="MP 16S fit to Preston log-normal model \n(pooled species)")

veiledspec(tb16_tax_occur_ss16509_prestonfit)
```

\
When computing Prestons' lognormal model fit without pooling data into groups, we miss ~251 species:

```{r preston_model2, echo=4}
tb16_tax_occur_ss16509_dist_all<-prestondistr(colSums(tb16_tax_occur_min16509_t))
plot(tb16_tax_occur_ss16509_prestonfit, main="MP 16S fit to Preston log-normal model \n(red: pooled; blue: non-pooled)")
lines(tb16_tax_occur_ss16509_dist_all, line.col="blue3")

veiledspec(tb16_tax_occur_ss16509_dist_all)
```


### 1.3.5) Rarefaction curves of rarefied and non-rarefied datasets

```{r rarefraction_curve, echo=F}
rarec_input<-t(as.matrix(colSums(tb16_tax_occur_ss16509_no_cero)))

#tb16_rarecurve_step1000_10000<-rarecurve(rarec_input, step = 1000, 10000, xlab = "Sample size", ylab = "OTUs", label = TRUE, main="18S amplicons diversity step=1000 & ss=10000\n(40,816 OTUs; 5,247,375 reads)\n")
#tb16_rarecurve_step1000_10000

tb16_rarecurve_step10_40000<-rarecurve(rarec_input, step = 1000, xlab = "Sample size", ylab = "OTUs", label = TRUE, main="16S amplicons diversity, step=1000 \n(8,173 OTUs; 1,502,319 reads)\n")

#without rarefying
rarec_allOTUs_input<-t(as.matrix(colSums(t(tb16_tax_occur))))
tb16_rarecurve_allOTUs_step1000<-rarecurve(rarec_allOTUs_input, step = 1000, xlab = "Sample size", ylab = "OTUs", label = TRUE, main="16S amplicons diversity non-rarefied step=1000 \n(9,114 OTUs; 5,604,699 reads)\n")
```


### 1.3.6) Beta diversity

#### 1.3.6.1) Dissimilarity matrix using Bray-Curtis index:

The Bray-Curtis dissimilarity, constrained between 0 (minimum distance) and 1 (highest dissimilarity) allows us to quantify the differences between samples according to the composition and relative abundance of their OTUs. In our dataset, most of the samples pairs take dissimilarity values between between 2 and 4, meaning that their composition is substantially similar.

```{r beta_div1, echo=FALSE}
#?vegdist
tb16_tax_occur_ss16509_no_cero.bray<-vegdist(tb16_tax_occur_ss16509_no_cero, method="bray")
boxplot(tb16_tax_occur_ss16509_no_cero.bray, main="Bray-Curtis dissimilarity matrix")
```

#### 1.3.6.2) Hierarchical clustering

The stations seem to form clusters according to geographic localization, but there are no evident clusters separated from the general groups.

(To be done: assign Longhurst provinces information to each station and check if any of the central clusters is meaningful regarding to the samples' geographical location)

```{r beta_div2, echo=FALSE, results="hide"}
#UPGMA
tb16_tax_occur_ss16509_no_cero.upgma<-hclust(tb16_tax_occur_ss16509_no_cero.bray, "average")
plot(tb16_tax_occur_ss16509_no_cero.upgma, cex=.35, main="Samples Hierarchical Clustering")
```

#### 1.3.6.3) Non-metric multidimensional scaling

We can identify a prominent group in the central part of the NMDS plot and a few outliers in the central-high edge of the plot. The stress parameter takes a value below 0.2, suggesting that the plot is acceptable. 

```{r monoNMDS, echo=F}
#NMDS
tb16_tax_occur_ss16509_no_cero.nmds<-monoMDS(tb16_tax_occur_ss16509_no_cero.bray)
tb16_tax_occur_ss16509_no_cero.nmds
plot(tb16_tax_occur_ss16509_no_cero.nmds, main="monoMDs method")
```


```{r metaNMDS, echo=F, results="hide", warning=F}
#When implementing a most robut function for computing NMDS plots, the result is quiet the same:

tb16_tax_occur_ss16509_no_cero.meta_nmds<-metaMDS(tb16_tax_occur_ss16509_no_cero.bray)
plot(tb16_tax_occur_ss16509_no_cero.meta_nmds, main="metaMDS method")
```

## 1.4) Geographical analysis

```{r load_geo_data, echo=F, results="hide", message=F, warning=F}
#load geographical ubication of stations and sort according to otu_tb16 stations sequence.
MP_geo_18S<-read.table(file="~/Documents/TFM2/data/MALASPINA/geo/mp_surface_ubication_modif.txt", sep="\t", header=T)

MP_geo_18S<-data.table(MP_geo_18S)
MP_geo_18S[,st_name:=(paste("st",MP_geo_18S$station,"_MD",MP_geo_18S$code,sep=""))]
MP_geo_18S<-MP_geo_18S[mixedorder(MP_geo_18S$st_name),]
tb16_tax_occur_ss16509_no_cero<-tb16_tax_occur_ss16509_no_cero[order(row.names(tb16_tax_occur_ss16509_no_cero)),]
#select only stations apearing in tb16_tax_occur_ss16509_no_cero:
MP_geo_18S<-MP_geo_18S[MP_geo_18S$st_name %in% row.names(tb16_tax_occur_ss16509_no_cero),]
row.names(MP_geo_18S)<-MP_geo_18S$st_name

dim(MP_geo_18S)
MP_geo_18S[1:5,1:5]
tb16_tax_occur_ss16509_no_cero[1:5,1:5]

#we'll use a function from library(fossil) to read lat-long in decimal degrees and translate into distance in km. We'll print only columns containing info about station, latitude and longitude:
MP_geo_18S_reduced<-create.lats(MP_geo_18S, loc="st_name", long="long", lat="lat")
head(MP_geo_18S_reduced)

#create a distance matrix (lower triangle) between a list of points.
geo_distances_MP_18S<-earth.dist(MP_geo_18S_reduced, dist = TRUE)
head(geo_distances_MP_18S)
dim(geo_distances_MP_18S)

geo_distances_MP_18S<-as.matrix(geo_distances_MP_18S)
dim(geo_distances_MP_18S)

#geo distances dataset ready to use: "geo_distances_MP_18S"
```

```{r working_datasets1, echo=F, results="hide", warning=F}
#working datasets:

#1) Community matrix: tb16_tax_occur_ss16509_no_cero
dim(tb16_tax_occur_ss16509_no_cero)
tb16_tax_occur_ss16509_no_cero[1:5, 1:5]

#2) Community Bray-Curtis: tb16_tax_occur_ss16509_no_cero.bray
tb16_tax_occur_ss16509_no_cero.bray<-as.matrix(tb16_tax_occur_ss16509_no_cero.bray)
dim(tb16_tax_occur_ss16509_no_cero.bray)

#3) Stations distances in km: geo_distances_MP_18S
dim(geo_distances_MP_18S)
```

Communities quickly change their composition across geographical distances:

```{r working_datasets4, echo=F}
plot(geo_distances_MP_18S, tb16_tax_occur_ss16509_no_cero.bray, pch=19, cex=0.4, xlab="Geopgraphical distances", ylab="Bray-Curtis dissimilarities")
```

### 1.4.1) Mantel correlograms

Mantel statistic is -significantlly- so low, meaning that the correlation between samples dissimilarity and geographical distances is weak.

```{r mantel_correlogram1, echo=F}
mantel(geo_distances_MP_18S, tb16_tax_occur_ss16509_no_cero.bray)
```


```{r mantel_correlogram2, echo=F, results="hide"}
#Maximum distance between samples:
max(geo_distances_MP_18S)

#Minimum distance between samples:
min(geo_distances_MP_18S)
```

Correlograms:

```{r mantel_correlogram4, echo=T}
MP_18s_ss16509_mantel_correl_by_1000km<-mantel.correlog(tb16_tax_occur_ss16509_no_cero.bray, D.geo=geo_distances_MP_18S, break.pts=seq(0,20000, by=1000))
plot(MP_18s_ss16509_mantel_correl_by_1000km)

MP_18s_ss16509_mantel_correl_by_100km<-mantel.correlog(tb16_tax_occur_ss16509_no_cero.bray, D.geo=geo_distances_MP_18S, break.pts=seq(0,20000, by=100))
plot(MP_18s_ss16509_mantel_correl_by_100km)
```

## 1.5) Abundance vs. occurence

```{r OTUs_mean_relative_abund, echo=F, results="hide"}
tb16_tax_occur_ss16509_no_cero[1:5,1:5]
tb16_tax_occur_ss16509_no_cero_t<-t(tb16_tax_occur_ss16509_no_cero)

colSums(tb16_tax_occur_ss16509_no_cero_t)

#local abundance percentage
tb16_tax_occur_ss16509_no_cero_t.rabund<-tb16_tax_occur_ss16509_no_cero_t/16509

colSums(tb16_tax_occur_ss16509_no_cero_t.rabund)
tb16_tax_occur_ss16509_no_cero_t.rabund[1:5,1:5]

#OTUs mean relative abundance
tb16_tax_occur_ss16509_no_cero_t.rabund_means<-rowMeans(tb16_tax_occur_ss16509_no_cero_t.rabund) 
tb16_tax_occur_ss16509_no_cero_t.rabund_means<-as.data.frame(tb16_tax_occur_ss16509_no_cero_t.rabund_means)

head(tb16_tax_occur_ss16509_no_cero_t.rabund_means)
```

```{r OTUs_occurence, echo=F, results='hide'}
tb16_tax_occur_ss16509_no_cero_t.rabund.occur<-tb16_tax_occur_ss16509_no_cero_t.rabund
tb16_tax_occur_ss16509_no_cero_t.rabund.occur[tb16_tax_occur_ss16509_no_cero_t.rabund.occur>0]<-1
tb16_tax_occur_ss16509_no_cero_t.rabund.occur[1:5,1:5] # presence - absence table

#percentage of occurence in overall stations
tb16_tax_occur_ss16509_no_cero_t.rabund_means.occurence_perc<-as.data.frame(100*(rowSums(tb16_tax_occur_ss16509_no_cero_t.rabund.occur)/91))

str(tb16_tax_occur_ss16509_no_cero_t.rabund_means.occurence_perc)
```

```{r merge_rabund_peroccur, echo=F, results='hide'}
otu_tb16_ss16509_rabund_percoccur<-merge(tb16_tax_occur_ss16509_no_cero_t.rabund_means,tb16_tax_occur_ss16509_no_cero_t.rabund_means.occurence_perc, by="row.names")

colnames(otu_tb16_ss16509_rabund_percoccur)<-c("OTUs","mean_rabund","perc_occur")
otu_tb16_ss16509_rabund_percoccur[1:5,]

row.names(otu_tb16_ss16509_rabund_percoccur)<-otu_tb16_ss16509_rabund_percoccur[,1]
otu_tb16_ss16509_rabund_percoccur<-otu_tb16_ss16509_rabund_percoccur[,-1]
otu_tb16_ss16509_rabund_percoccur[1:5,]
```

OTUs distribution according to their percentage of occurence and relative abundance.\
- red line: OTUs that occur in more than 80% of the samples.\
- blue line: regionally abundant OTUs (> 0.1%).\
- green line: regionally rare (< 0.001%).

```{r abund_vs_occurence_table, echo=F}
plot(otu_tb16_ss16509_rabund_percoccur$mean_rabund,otu_tb16_ss16509_rabund_percoccur$perc_occur, log="x", pch=19, cex=0.8, xlab="Mean relative abundance", ylab="Percentage of occurence")
abline(h=80, col="red") #occurence higher than 80%
abline(v=0.00001, col="green") #rare OTUs
abline(v=0.001, col="blue") #cosmopolitan OTUs

#Conventional limits:
#Regionally rare     = 0.00001
#Regionally abundant = 0.001
```

Regionally abundant OTUs (relative abundance over 0.1%):

```{r abundant_OTUs, echo=F}
#regionally abundant
tb16_ss16509_abundant<-otu_tb16_ss16509_rabund_percoccur[otu_tb16_ss16509_rabund_percoccur$mean_rabund > 0.001,]

tb16_ss16509_abundant_sorted<-tb16_ss16509_abundant[order(tb16_ss16509_abundant$mean_rabund, tb16_ss16509_abundant$perc_occur, decreasing = T), c(1,2)]

tb16_ss16509_abundant_sorted_prov<-cbind(otu_names=row.names(tb16_ss16509_abundant_sorted),tb16_ss16509_abundant_sorted)
tb16_class_prov<-cbind(otu_names=row.names(tb16_class),tb16_class)
#tb16_class_prov<-tb16_class_prov[,-2]
tb16_ss16509_abundant_sorted_prov<-merge(tb16_ss16509_abundant_sorted_prov,tb16_class_prov, by="otu_names", all.x=TRUE)

tb16_ss16509_abundant_sorted_prov<-tb16_ss16509_abundant_sorted_prov[order( tb16_ss16509_abundant_sorted_prov$mean_rabund, tb16_ss16509_abundant_sorted_prov$perc_occur, decreasing = T), c(1,2,3,9,4,6)]
tb16_ss16509_abundant_sorted_prov
#dim(tb16_ss16509_abundant_sorted_prov)
```

Number and roportion of regionally abundant OTUs (%):

```{r abundant_OTUs2, echo=F}
nrow(tb16_ss16509_abundant_sorted_prov)

(nrow(tb16_ss16509_abundant_sorted_prov)/nrow(otu_tb16_ss16509_rabund_percoccur))*100
```

\
Cosmopolitan OTUs (relative abundance over 0.1% and occurence in more than 80% of samples):

```{r select_cosmopolitan, echo=F}
otu_tb16_ss16509_rabund_cosm<-otu_tb16_ss16509_rabund_percoccur[otu_tb16_ss16509_rabund_percoccur$mean_rabund > 0.001,]
otu_tb16_ss16509_rabund_poccur_cosm<-otu_tb16_ss16509_rabund_cosm[otu_tb16_ss16509_rabund_cosm$perc_occur > 80,]
otu_tb16_ss16509_cosmop_sorted<-otu_tb16_ss16509_rabund_poccur_cosm[order(otu_tb16_ss16509_rabund_poccur_cosm$perc_occur, otu_tb16_ss16509_rabund_poccur_cosm$mean_rabund, decreasing = T), c(1,2)]


otu_tb16_ss16509_cosmop_sorted_prov<-cbind(otu_names=row.names(otu_tb16_ss16509_cosmop_sorted),otu_tb16_ss16509_cosmop_sorted)
otu_tb16_ss16509_cosmop_sorted_prov<-merge(otu_tb16_ss16509_cosmop_sorted_prov,tb16_class_prov, by="otu_names", all.x=TRUE)

otu_tb16_ss16509_cosmop_sorted_prov<-otu_tb16_ss16509_cosmop_sorted_prov[order(otu_tb16_ss16509_cosmop_sorted_prov$perc_occur, otu_tb16_ss16509_cosmop_sorted_prov$mean_rabund, decreasing = T), c(1,2,3,9,4,6)]
otu_tb16_ss16509_cosmop_sorted_prov
#dim(otu_tb16_ss16509_cosmop_sorted)
```

Number and proportion (%) of cosmopolitan OTUs:

```{r percentage_cosmopolitan, echo=F}
nrow(otu_tb16_ss16509_cosmop_sorted)

#otu_tb16_ss16509_cosmop_sorted_prov<-cbind(otu_names=row.names(otu_tb16_ss16509_cosmop_sorted),otu_tb16_ss16509_cosmop_sorted)
#otu_tb16_ss16509_cosmop_sorted_prov<-merge(otu_tb16_ss16509_cosmop_sorted_prov,tb16_class_prov, by="otu_names", all.x=TRUE)

(nrow(otu_tb16_ss16509_cosmop_sorted)/nrow(otu_tb16_ss16509_rabund_percoccur))*100
```

Number and proportion (%) of rare OTUs:

```{r rare_OTUs, echo=F}
nrow(otu_tb16_ss16509_rabund_percoccur[otu_tb16_ss16509_rabund_percoccur$mean_rabund < 0.00001 & otu_tb16_ss16509_rabund_percoccur$mean_rabund >0,])

(nrow(otu_tb16_ss16509_rabund_percoccur[otu_tb16_ss16509_rabund_percoccur$mean_rabund < 0.00001 & otu_tb16_ss16509_rabund_percoccur$mean_rabund >0,])/nrow(otu_tb16_ss16509_rabund_percoccur))*100
```



## 2.6) Taxonomic composition analysis

### 2.6.1) Normalized data

```{r merge_tables_16S, echo=FALSE, results="hide", warning=FALSE}
#Let's add the taxonomic classification by merging "tb16_tax_occur_ss16509_no_cero" with "tb16_tax":

tb16_tax_occur_ss16509_no_cero_t<-t(tb16_tax_occur_ss16509_no_cero)
tb16_ss16509_tax<-merge(tb16_tax_occur_ss16509_no_cero_t,tb16_class, by="row.names", all.x=T)

dim(tb16_ss16509_tax)
tb16_ss16509_tax[1:5,1:5]

#fix OTU_no as new row
rownames(tb16_ss16509_tax)=tb16_ss16509_tax$Row.names
#add OTU_no as rowname
rownames.tb16_ss16509_tax<-tb16_ss16509_tax[,1]
tb16_ss16509_tax<-tb16_ss16509_tax[,-1]

dim(tb16_ss16509_tax)
tb16_ss16509_tax[1:5, 1:5]

#sort by OTU_no (split rowname, introduce no. into column "OTU_no" and sort)
tb16_ss16509_tax["OTU_no"] <- NA
tb16_ss16509_tax$OTU_no <- sapply(strsplit(rownames(tb16_ss16509_tax),split= "\\_"),'[',2)
tb16_ss16509_tax$OTU_no <- as.numeric(as.character(tb16_ss16509_tax$OTU_no))
tb16_ss16509_tax_sorted<-tb16_ss16509_tax[order(tb16_ss16509_tax$OTU_no, decreasing = FALSE), ]
```

No. of OTUs and reads of the rearefied dataset:

```{r dim_merge1_16S, echo=F}
nrow(tb16_ss16509_tax_sorted)
sum(colSums(tb16_ss16509_tax_sorted[,1:91]))

#dim(tb16_ss16509_tax_sorted)
#tb16_ss16509_tax_sorted[1:5,1:5]
```

\
Number and proportion (%) of OTUs of phototrophic groups:

```{r dim_merge2_v2, echo=F}
#selection of phototrophs:
tb16_phototrophs <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A != "other_bacteria"),]

nrow(tb16_phototrophs)
nrow(tb16_phototrophs)*100/nrow(tb16_ss16509_tax_sorted)
```

Number and proportion (%) of reads of phototrophic groups:

```{r dim_merge2.2_v2, echo=F}
sum(colSums(tb16_phototrophs[,1:91]))
(sum(colSums(tb16_phototrophs[,1:91])))*100/sum(colSums(tb16_ss16509_tax_sorted[,1:91]))

#dim(tb16_phototrophs)
#tb16_phototrophs[1:5,1:5]
```

\
Number and proportion (%) of OTUs of non-phototrophic groups:

```{r dim_merge3_v2, echo=F}
#selection of non.phototrophs:
tb16_non.phototrophs <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "other_bacteria"),]

nrow(tb16_non.phototrophs)
(nrow(tb16_non.phototrophs))*100/nrow(tb16_ss16509_tax_sorted)
```

Number and proportion (%) of reads of non-phototrophic groups:

```{r dim_merge3.2_v2, echo=F}
sum(colSums(tb16_non.phototrophs[,1:91]))
(sum(colSums(tb16_non.phototrophs[,1:91])))*100/sum(colSums(tb16_ss16509_tax_sorted[,1:91]))
```

\
**PHOTOTROPHS + HETEROTROPHS**

**Absolute values**


```{r count_samples_per_group_Heterotrophs_16S, echo=F, warning=FALSE, results="hide"}
occurrence_counts_phototrophs<-data.table()

#create a table per group and count in how many samples they occur. 
Dinophyceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Dinophyceae"),]
Dinophyceae_tb_occur <- Dinophyceae_tb[,1:91]
Dinophyceae_tb_occur_len<-length(Dinophyceae_tb_occur[,colSums(Dinophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Dinophyceae",samples_per_class=Dinophyceae_tb_occur_len))

Prasinophyceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "other_Prasinophyceae"),]
Prasinophyceae_tb_occur <- Prasinophyceae_tb[,1:91]
Prasinophyceae_tb_occur_len<-length(Prasinophyceae_tb_occur[,colSums(Prasinophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="other_Prasinophyceae",samples_per_class=Prasinophyceae_tb_occur_len))

Chrysophyceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Chrysophyceae"),]
Chrysophyceae_tb_occur <- Chrysophyceae_tb[,1:91]
Chrysophyceae_tb_occur_len<-length(Chrysophyceae_tb_occur[,colSums(Chrysophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Chrysophyceae",samples_per_class=Chrysophyceae_tb_occur_len))

Pelagophyceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Pelagophyceae"),]
Pelagophyceae_tb_occur <- Pelagophyceae_tb[,1:91]
Pelagophyceae_tb_occur_len<-length(Pelagophyceae_tb_occur[,colSums(Pelagophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Pelagophyceae",samples_per_class=Pelagophyceae_tb_occur_len))

Dictyochophyceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Dictyochophyceae"),]
Dictyochophyceae_tb_occur <- Dictyochophyceae_tb[,1:91]
Dictyochophyceae_tb_occur_len<-length(Dictyochophyceae_tb_occur[,colSums(Dictyochophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Dictyochophyceae",samples_per_class=Dictyochophyceae_tb_occur_len))

Cryptophyceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Cryptophyceae"),]
Cryptophyceae_tb_occur <- Cryptophyceae_tb[,1:91]
Cryptophyceae_tb_occur_len<-length(Cryptophyceae_tb_occur[,colSums(Cryptophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Cryptophyceae",samples_per_class=Cryptophyceae_tb_occur_len))

Bacillariophyceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Bacillariophyceae"),]
Bacillariophyceae_tb_occur <- Bacillariophyceae_tb[,1:91]
Bacillariophyceae_tb_occur_len<-length(Bacillariophyceae_tb_occur[,colSums(Bacillariophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Bacillariophyceae",samples_per_class=Bacillariophyceae_tb_occur_len))

Chlorarachniophyceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Chlorarachniophyceae"),]
Chlorarachniophyceae_tb_occur <- Chlorarachniophyceae_tb[,1:91]
Chlorarachniophyceae_tb_occur_len<-length(Chlorarachniophyceae_tb_occur[,colSums(Chlorarachniophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Chlorarachniophyceae",samples_per_class=Chlorarachniophyceae_tb_occur_len))

Bolidophyceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Bolidophyceae"),]
Bolidophyceae_tb_occur <- Bolidophyceae_tb[,1:91]
Bolidophyceae_tb_occur_len<-length(Bolidophyceae_tb_occur[,colSums(Bolidophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Bolidophyceae",samples_per_class=Bolidophyceae_tb_occur_len))

Pinguiophyceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Pinguiophyceae"),]
Pinguiophyceae_tb_occur <- Pinguiophyceae_tb[,1:91]
Pinguiophyceae_tb_occur_len<-length(Pinguiophyceae_tb_occur[,colSums(Pinguiophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Pinguiophyceae",samples_per_class=Pinguiophyceae_tb_occur_len))

Prymnesiophyceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Prymnesiophyceae"),]
Prymnesiophyceae_tb_occur <- Prymnesiophyceae_tb[,1:91]
Prymnesiophyceae_tb_occur_len<-length(Prymnesiophyceae_tb_occur[,colSums(Prymnesiophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Prymnesiophyceae",samples_per_class=Prymnesiophyceae_tb_occur_len))

Mamiellophyceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Mamiellophyceae"),]
Mamiellophyceae_tb_occur <- Mamiellophyceae_tb[,1:91]
Mamiellophyceae_tb_occur_len<-length(Mamiellophyceae_tb_occur[,colSums(Mamiellophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Mamiellophyceae",samples_per_class=Mamiellophyceae_tb_occur_len))

Eustigmatophyceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Eustigmatophyceae"),]
Eustigmatophyceae_tb_occur <- Eustigmatophyceae_tb[,1:91]
Eustigmatophyceae_tb_occur_len<-length(Eustigmatophyceae_tb_occur[,colSums(Eustigmatophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Eustigmatophyceae",samples_per_class=Eustigmatophyceae_tb_occur_len))

Chlorophyceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Chlorophyceae"),]
Chlorophyceae_tb_occur <- Chlorophyceae_tb[,1:91]
Chlorophyceae_tb_occur_len<-length(Chlorophyceae_tb_occur[,colSums(Chlorophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Chlorophyceae",samples_per_class=Chlorophyceae_tb_occur_len))

Ulvophyceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Ulvophyceae"),]
Ulvophyceae_tb_occur <- Ulvophyceae_tb[,1:91]
Ulvophyceae_tb_occur_len<-length(Ulvophyceae_tb_occur[,colSums(Ulvophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Ulvophyceae",samples_per_class=Ulvophyceae_tb_occur_len))

Raphydophyceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Raphydophyceae"),]
Raphydophyceae_tb_occur <- Raphydophyceae_tb[,1:91]
Raphydophyceae_tb_occur_len<-length(Raphydophyceae_tb_occur[,colSums(Raphydophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Raphydophyceae",samples_per_class=Raphydophyceae_tb_occur_len))

Trebouxiophyceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Trebouxiophyceae"),]
Trebouxiophyceae_tb_occur <- Trebouxiophyceae_tb[,1:91]
Trebouxiophyceae_tb_occur_len<-length(Trebouxiophyceae_tb_occur[,colSums(Trebouxiophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Trebouxiophyceae",samples_per_class=Trebouxiophyceae_tb_occur_len))

Phaeophyceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Phaeophyceae"),]
Phaeophyceae_tb_occur <- Phaeophyceae_tb[,1:91]
Phaeophyceae_tb_occur_len<-length(Phaeophyceae_tb_occur[,colSums(Phaeophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Phaeophyceae",samples_per_class=Phaeophyceae_tb_occur_len))

Phaeothamniophyceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Phaeothamniophyceae"),]
Phaeothamniophyceae_tb_occur <- Phaeothamniophyceae_tb[,1:91]
Phaeothamniophyceae_tb_occur_len<-length(Phaeothamniophyceae_tb_occur[,colSums(Phaeothamniophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Phaeothamniophyceae",samples_per_class=Phaeothamniophyceae_tb_occur_len))

Xanthophyceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Xanthophyceae"),]
Xanthophyceae_tb_occur <- Xanthophyceae_tb[,1:91]
Xanthophyceae_tb_occur_len<-length(Xanthophyceae_tb_occur[,colSums(Xanthophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Xanthophyceae",samples_per_class=Xanthophyceae_tb_occur_len))

Chlorodendrophyceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Chlorodendrophyceae"),]
Chlorodendrophyceae_tb_occur <- Chlorodendrophyceae_tb[,1:91]
Chlorodendrophyceae_tb_occur_len<-length(Chlorodendrophyceae_tb_occur[,colSums(Chlorodendrophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Chlorodendrophyceae",samples_per_class=Chlorodendrophyceae_tb_occur_len))

IncertaeSedis_Archaeplastida_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "IncertaeSedis_Archaeplastida"),]
IncertaeSedis_Archaeplastida_tb_occur <- IncertaeSedis_Archaeplastida_tb[,1:91]
IncertaeSedis_Archaeplastida_tb_occur_len<-length(IncertaeSedis_Archaeplastida_tb_occur[,colSums(IncertaeSedis_Archaeplastida_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="IncertaeSedis_Archaeplastida",samples_per_class=IncertaeSedis_Archaeplastida_tb_occur_len))

Nephroselmidophyceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Nephroselmidophyceae"),]
Nephroselmidophyceae_tb_occur <- Nephroselmidophyceae_tb[,1:91]
Nephroselmidophyceae_tb_occur_len<-length(Nephroselmidophyceae_tb_occur[,colSums(Nephroselmidophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Nephroselmidophyceae",samples_per_class=Nephroselmidophyceae_tb_occur_len))

Pavlovophyceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Pavlovophyceae"),]
Pavlovophyceae_tb_occur <- Pavlovophyceae_tb[,1:91]
Pavlovophyceae_tb_occur_len<-length(Pavlovophyceae_tb_occur[,colSums(Pavlovophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Pavlovophyceae",samples_per_class=Pavlovophyceae_tb_occur_len))

Rhodophyceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Rhodophyceae"),]
Rhodophyceae_tb_occur <- Rhodophyceae_tb[,1:91]
Rhodophyceae_tb_occur_len<-length(Rhodophyceae_tb_occur[,colSums(Rhodophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Rhodophyceae",samples_per_class=Rhodophyceae_tb_occur_len))

Rappemonads_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Rappemonads"),]
Rappemonads_tb_occur <- Rappemonads_tb[,1:91]
Rappemonads_tb_occur_len<-length(Rappemonads_tb_occur[,colSums(Rappemonads_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Rappemonads",samples_per_class=Rappemonads_tb_occur_len))

MOCH_1_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "MOCH_1"),]
MOCH_1_tb_occur <- MOCH_1_tb[,1:91]
MOCH_1_tb_occur_len<-length(MOCH_1_tb_occur[,colSums(MOCH_1_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="MOCH_1",samples_per_class=MOCH_1_tb_occur_len))

MOCH_2_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "MOCH_2"),]
MOCH_2_tb_occur <- MOCH_2_tb[,1:91]
MOCH_2_tb_occur_len<-length(MOCH_2_tb_occur[,colSums(MOCH_2_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="MOCH_2",samples_per_class=MOCH_2_tb_occur_len))

MOCH_5_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "MOCH_5"),]
MOCH_5_tb_occur <- MOCH_5_tb[,1:91]
MOCH_5_tb_occur_len<-length(MOCH_5_tb_occur[,colSums(MOCH_5_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="MOCH_5",samples_per_class=MOCH_5_tb_occur_len))

Prasinophyceae_clade_VII_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Prasinophyceae_clade-VII"),]
Prasinophyceae_clade_VII_tb_occur <- Prasinophyceae_clade_VII_tb[,1:91]
Prasinophyceae_clade_VII_tb_occur_len<-length(Prasinophyceae_clade_VII_tb_occur[,colSums(Prasinophyceae_clade_VII_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Prasinophyceae_clade-VII",samples_per_class=Prasinophyceae_clade_VII_tb_occur_len))

Prasinophyceae_clade_IX_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Prasinophyceae_clade-IX"),]
Prasinophyceae_clade_IX_tb_occur <- Prasinophyceae_clade_IX_tb[,1:91]
Prasinophyceae_clade_IX_tb_occur_len<-length(Prasinophyceae_clade_IX_tb_occur[,colSums(Prasinophyceae_clade_IX_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Prasinophyceae_clade-IX",samples_per_class=Prasinophyceae_clade_IX_tb_occur_len))

Pyramimonadaceae_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Pyramimonadaceae"),]
Pyramimonadaceae_tb_occur <- Pyramimonadaceae_tb[,1:91]
Pyramimonadaceae_tb_occur_len<-length(Pyramimonadaceae_tb_occur[,colSums(Pyramimonadaceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Pyramimonadaceae",samples_per_class=Pyramimonadaceae_tb_occur_len))

Cyanobacteria_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Cyanobacteria"),]
Cyanobacteria_tb_occur <- Cyanobacteria_tb[,1:91]
Cyanobacteria_tb_occur_len<-length(Cyanobacteria_tb_occur[,colSums(Cyanobacteria_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Cyanobacteria",samples_per_class=Cyanobacteria_tb_occur_len))

other_cyanobacteria_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "other_cyanobacteria"),]
other_cyanobacteria_tb_occur <- other_cyanobacteria_tb[,1:91]
other_cyanobacteria_tb_occur_len<-length(other_cyanobacteria_tb_occur[,colSums(other_cyanobacteria_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="other_cyanobacteria",samples_per_class=other_cyanobacteria_tb_occur_len))

other_bacteria_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "other_bacteria"),]
other_bacteria_tb_occur <- other_bacteria_tb[,1:91]
other_bacteria_tb_occur_len<-length(other_bacteria_tb_occur[,colSums(other_bacteria_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="other_bacteria",samples_per_class=other_bacteria_tb_occur_len))

other_plastids_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "other_plastids"),]
other_plastids_tb_occur <- other_plastids_tb[,1:91]
other_plastids_tb_occur_len<-length(other_plastids_tb_occur[,colSums(other_plastids_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="other_plastids",samples_per_class=other_plastids_tb_occur_len))

occurrence_counts_phototrophs
#row.names(occurrence_counts_phototrophs)<-occurrence_counts_phototrophs$group
occurrence_counts_phototrophs<-as.data.frame(occurrence_counts_phototrophs)
```

```{r aggregate_Heterotrophs_16S, echo=FALSE, results="hide", warning=FALSE}
class_summary_reads_per_class<-aggregate(rowSums(tb16_ss16509_tax_sorted[1:91]), list(tb16_ss16509_tax_sorted$class_A), sum)
# count the different groups

class_summary_otus_per_class<-aggregate(rowSums(tb16_ss16509_tax_sorted[1:91]), list(tb16_ss16509_tax_sorted$class_A), length)

## READS PER CLASS ##
attach(class_summary_reads_per_class)
class_summary_reads_per_class_order<-class_summary_reads_per_class[order(-x),]
detach(class_summary_reads_per_class)
class_summary_reads_per_class_order

#fix column names
row.names(class_summary_reads_per_class_order)<-class_summary_reads_per_class_order[,1]
class_summary_reads_per_class_order<-class_summary_reads_per_class_order[c(-1)]
colnames(class_summary_reads_per_class_order)<-c("reads_per_class")

## OTUs PER CLASS ##
attach(class_summary_otus_per_class)
class_summary_otus_per_class_order<-class_summary_otus_per_class[order(-x),]
detach(class_summary_otus_per_class)
class_summary_otus_per_class_order

row.names(class_summary_otus_per_class_order)<-class_summary_otus_per_class_order[,1]
class_summary_otus_per_class_order<-class_summary_otus_per_class_order[c(-1)]
colnames(class_summary_otus_per_class_order)<-c("OTUs_per_class")

#class_summary_reads<-aggregate(sum~class, data=otutab_full_wTax, FUN="sum") 
# sum reads different groups
```

```{r create_table_merging_#OTUs_#reads_#samples_Heterotrophs_16S, echo=FALSE, warning=F}
tb16S_OTUs_reads_samples <- merge(class_summary_reads_per_class_order, class_summary_otus_per_class_order, by="row.names")

row.names(tb16S_OTUs_reads_samples)<-tb16S_OTUs_reads_samples[,1]

tb16S_OTUs_reads_samples<-tb16S_OTUs_reads_samples[,-1]
colnames(tb16S_OTUs_reads_samples)<-c("reads_per_class","OTUs_per_class")
#tb16S_OTUs_reads_samples[1:5,1:2]

tb16S_OTUs_reads_samples<-tb16S_OTUs_reads_samples[order(tb16S_OTUs_reads_samples$reads_per_class, tb16S_OTUs_reads_samples$OTUs_per_class, decreasing = T), c(1,2)]

#add no. of samples per class
tb16S_OTUs_reads_samples<-merge(tb16S_OTUs_reads_samples, occurrence_counts_phototrophs, by.x="row.names", by.y="group", all.x=TRUE)

row.names(tb16S_OTUs_reads_samples)<-tb16S_OTUs_reads_samples[,1]
tb16S_OTUs_reads_samples<-tb16S_OTUs_reads_samples[,-1]
tb16S_OTUs_reads_samples
```


\
**Relative values**

```{r OTUs_vs_reads_relative_values_16S_Heterotrophs, echo=FALSE, warning=F}

tb16S_OTUs_reads_samples_rel_abund <- tb16S_OTUs_reads_samples

tb16S_OTUs_reads_samples_rel_abund$reads_per_class<-(tb16S_OTUs_reads_samples_rel_abund$reads_per_class*100)/colSums(tb16S_OTUs_reads_samples)[1]
tb16S_OTUs_reads_samples_rel_abund$OTUs_per_class<-(tb16S_OTUs_reads_samples_rel_abund$OTUs_per_class*100)/colSums(tb16S_OTUs_reads_samples)[2]
tb16S_OTUs_reads_samples_rel_abund$samples_per_class<-(tb16S_OTUs_reads_samples_rel_abund$samples_per_class*100)/91

colSums(tb16S_OTUs_reads_samples_rel_abund)
tb16S_OTUs_reads_samples_rel_abund
```

\
\

Reads per class vs. OTUs per class:

```{r OTUs_vs_reads_16S_v2_Heterotrophs, echo=FALSE, warnings=F, warning=F}
ggplot(tb16S_OTUs_reads_samples_rel_abund) +
  geom_point(aes(OTUs_per_class, reads_per_class)) +
  geom_text_repel(aes(OTUs_per_class, reads_per_class, label = rownames(tb16S_OTUs_reads_samples)), size=4, force = 3, box.padding = unit(0.1, "lines"), point.padding = unit(0.02, 'lines')) + labs(title="[B1] MP 16S: reads per class vs. OTUs per class", x="\ndiversity (no. of OTUs in %)", y="abundance (no. of reads in %)\n") + scale_x_log10(limits = c(-1,200), breaks = c(0,0.1,1,10,100)) + scale_y_log10(limits = c(-1,200), breaks = c(0,0.1,1,10,100))  + theme(axis.text=element_text(size=14), axis.title=element_text(size=16))
```

\
\

Reads per class vs. samples in which they occurr:

```{r reads_vs_samples_16S_Heterotrophs, echo=FALSE, warning=F}
ggplot(tb16S_OTUs_reads_samples_rel_abund) +
  geom_point(aes(samples_per_class,reads_per_class)) +
  geom_text_repel(aes(samples_per_class,reads_per_class, label = rownames(tb16S_OTUs_reads_samples)), size=4, force = 8, box.padding = unit(0.1, "lines"), point.padding = unit(0.2, 'lines')) + scale_x_continuous(limits = c(0, 116), breaks=c(25,50,75,100)) + labs(title="[B2] MP 16S: reads per class vs. samples in which they occur", x="\nsamples in which they occurr (%)", y="abundance (no. of reads in %)\n") + scale_y_log10(limits = c(-1,200), breaks = c(0,0.1,1,10,100)) + theme(axis.text=element_text(size=14), axis.title=element_text(size=16))
```



\
**PHOTOTROPHS**

```{r count_samples_per_group_16S, echo=F, warning=FALSE, results="hide"}
occurrence_counts_phototrophs<-data.table()

#create a table per group and count in how many samples they occur. 
Dinophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Dinophyceae"),]
Dinophyceae_tb_occur <- Dinophyceae_tb[,1:91]
Dinophyceae_tb_occur_len<-length(Dinophyceae_tb_occur[,colSums(Dinophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Dinophyceae",samples_per_class=Dinophyceae_tb_occur_len))

Prasinophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "other_Prasinophyceae"),]
Prasinophyceae_tb_occur <- Prasinophyceae_tb[,1:91]
Prasinophyceae_tb_occur_len<-length(Prasinophyceae_tb_occur[,colSums(Prasinophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="other_Prasinophyceae",samples_per_class=Prasinophyceae_tb_occur_len))

Chrysophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Chrysophyceae"),]
Chrysophyceae_tb_occur <- Chrysophyceae_tb[,1:91]
Chrysophyceae_tb_occur_len<-length(Chrysophyceae_tb_occur[,colSums(Chrysophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Chrysophyceae",samples_per_class=Chrysophyceae_tb_occur_len))

Pelagophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Pelagophyceae"),]
Pelagophyceae_tb_occur <- Pelagophyceae_tb[,1:91]
Pelagophyceae_tb_occur_len<-length(Pelagophyceae_tb_occur[,colSums(Pelagophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Pelagophyceae",samples_per_class=Pelagophyceae_tb_occur_len))

Dictyochophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Dictyochophyceae"),]
Dictyochophyceae_tb_occur <- Dictyochophyceae_tb[,1:91]
Dictyochophyceae_tb_occur_len<-length(Dictyochophyceae_tb_occur[,colSums(Dictyochophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Dictyochophyceae",samples_per_class=Dictyochophyceae_tb_occur_len))

Cryptophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Cryptophyceae"),]
Cryptophyceae_tb_occur <- Cryptophyceae_tb[,1:91]
Cryptophyceae_tb_occur_len<-length(Cryptophyceae_tb_occur[,colSums(Cryptophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Cryptophyceae",samples_per_class=Cryptophyceae_tb_occur_len))

Bacillariophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Bacillariophyceae"),]
Bacillariophyceae_tb_occur <- Bacillariophyceae_tb[,1:91]
Bacillariophyceae_tb_occur_len<-length(Bacillariophyceae_tb_occur[,colSums(Bacillariophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Bacillariophyceae",samples_per_class=Bacillariophyceae_tb_occur_len))

Chlorarachniophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Chlorarachniophyceae"),]
Chlorarachniophyceae_tb_occur <- Chlorarachniophyceae_tb[,1:91]
Chlorarachniophyceae_tb_occur_len<-length(Chlorarachniophyceae_tb_occur[,colSums(Chlorarachniophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Chlorarachniophyceae",samples_per_class=Chlorarachniophyceae_tb_occur_len))

Bolidophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Bolidophyceae"),]
Bolidophyceae_tb_occur <- Bolidophyceae_tb[,1:91]
Bolidophyceae_tb_occur_len<-length(Bolidophyceae_tb_occur[,colSums(Bolidophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Bolidophyceae",samples_per_class=Bolidophyceae_tb_occur_len))

Pinguiophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Pinguiophyceae"),]
Pinguiophyceae_tb_occur <- Pinguiophyceae_tb[,1:91]
Pinguiophyceae_tb_occur_len<-length(Pinguiophyceae_tb_occur[,colSums(Pinguiophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Pinguiophyceae",samples_per_class=Pinguiophyceae_tb_occur_len))

Prymnesiophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Prymnesiophyceae"),]
Prymnesiophyceae_tb_occur <- Prymnesiophyceae_tb[,1:91]
Prymnesiophyceae_tb_occur_len<-length(Prymnesiophyceae_tb_occur[,colSums(Prymnesiophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Prymnesiophyceae",samples_per_class=Prymnesiophyceae_tb_occur_len))

Mamiellophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Mamiellophyceae"),]
Mamiellophyceae_tb_occur <- Mamiellophyceae_tb[,1:91]
Mamiellophyceae_tb_occur_len<-length(Mamiellophyceae_tb_occur[,colSums(Mamiellophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Mamiellophyceae",samples_per_class=Mamiellophyceae_tb_occur_len))

Eustigmatophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Eustigmatophyceae"),]
Eustigmatophyceae_tb_occur <- Eustigmatophyceae_tb[,1:91]
Eustigmatophyceae_tb_occur_len<-length(Eustigmatophyceae_tb_occur[,colSums(Eustigmatophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Eustigmatophyceae",samples_per_class=Eustigmatophyceae_tb_occur_len))

Chlorophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Chlorophyceae"),]
Chlorophyceae_tb_occur <- Chlorophyceae_tb[,1:91]
Chlorophyceae_tb_occur_len<-length(Chlorophyceae_tb_occur[,colSums(Chlorophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Chlorophyceae",samples_per_class=Chlorophyceae_tb_occur_len))

Ulvophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Ulvophyceae"),]
Ulvophyceae_tb_occur <- Ulvophyceae_tb[,1:91]
Ulvophyceae_tb_occur_len<-length(Ulvophyceae_tb_occur[,colSums(Ulvophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Ulvophyceae",samples_per_class=Ulvophyceae_tb_occur_len))

Raphydophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Raphydophyceae"),]
Raphydophyceae_tb_occur <- Raphydophyceae_tb[,1:91]
Raphydophyceae_tb_occur_len<-length(Raphydophyceae_tb_occur[,colSums(Raphydophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Raphydophyceae",samples_per_class=Raphydophyceae_tb_occur_len))

Trebouxiophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Trebouxiophyceae"),]
Trebouxiophyceae_tb_occur <- Trebouxiophyceae_tb[,1:91]
Trebouxiophyceae_tb_occur_len<-length(Trebouxiophyceae_tb_occur[,colSums(Trebouxiophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Trebouxiophyceae",samples_per_class=Trebouxiophyceae_tb_occur_len))

Phaeophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Phaeophyceae"),]
Phaeophyceae_tb_occur <- Phaeophyceae_tb[,1:91]
Phaeophyceae_tb_occur_len<-length(Phaeophyceae_tb_occur[,colSums(Phaeophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Phaeophyceae",samples_per_class=Phaeophyceae_tb_occur_len))

Phaeothamniophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Phaeothamniophyceae"),]
Phaeothamniophyceae_tb_occur <- Phaeothamniophyceae_tb[,1:91]
Phaeothamniophyceae_tb_occur_len<-length(Phaeothamniophyceae_tb_occur[,colSums(Phaeothamniophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Phaeothamniophyceae",samples_per_class=Phaeothamniophyceae_tb_occur_len))

Xanthophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Xanthophyceae"),]
Xanthophyceae_tb_occur <- Xanthophyceae_tb[,1:91]
Xanthophyceae_tb_occur_len<-length(Xanthophyceae_tb_occur[,colSums(Xanthophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Xanthophyceae",samples_per_class=Xanthophyceae_tb_occur_len))

Chlorodendrophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Chlorodendrophyceae"),]
Chlorodendrophyceae_tb_occur <- Chlorodendrophyceae_tb[,1:91]
Chlorodendrophyceae_tb_occur_len<-length(Chlorodendrophyceae_tb_occur[,colSums(Chlorodendrophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Chlorodendrophyceae",samples_per_class=Chlorodendrophyceae_tb_occur_len))

IncertaeSedis_Archaeplastida_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "IncertaeSedis_Archaeplastida"),]
IncertaeSedis_Archaeplastida_tb_occur <- IncertaeSedis_Archaeplastida_tb[,1:91]
IncertaeSedis_Archaeplastida_tb_occur_len<-length(IncertaeSedis_Archaeplastida_tb_occur[,colSums(IncertaeSedis_Archaeplastida_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="IncertaeSedis_Archaeplastida",samples_per_class=IncertaeSedis_Archaeplastida_tb_occur_len))

Nephroselmidophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Nephroselmidophyceae"),]
Nephroselmidophyceae_tb_occur <- Nephroselmidophyceae_tb[,1:91]
Nephroselmidophyceae_tb_occur_len<-length(Nephroselmidophyceae_tb_occur[,colSums(Nephroselmidophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Nephroselmidophyceae",samples_per_class=Nephroselmidophyceae_tb_occur_len))

Pavlovophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Pavlovophyceae"),]
Pavlovophyceae_tb_occur <- Pavlovophyceae_tb[,1:91]
Pavlovophyceae_tb_occur_len<-length(Pavlovophyceae_tb_occur[,colSums(Pavlovophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Pavlovophyceae",samples_per_class=Pavlovophyceae_tb_occur_len))

Rhodophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Rhodophyceae"),]
Rhodophyceae_tb_occur <- Rhodophyceae_tb[,1:91]
Rhodophyceae_tb_occur_len<-length(Rhodophyceae_tb_occur[,colSums(Rhodophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Rhodophyceae",samples_per_class=Rhodophyceae_tb_occur_len))

Rappemonads_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Rappemonads"),]
Rappemonads_tb_occur <- Rappemonads_tb[,1:91]
Rappemonads_tb_occur_len<-length(Rappemonads_tb_occur[,colSums(Rappemonads_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Rappemonads",samples_per_class=Rappemonads_tb_occur_len))

MOCH_1_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "MOCH_1"),]
MOCH_1_tb_occur <- MOCH_1_tb[,1:91]
MOCH_1_tb_occur_len<-length(MOCH_1_tb_occur[,colSums(MOCH_1_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="MOCH_1",samples_per_class=MOCH_1_tb_occur_len))

MOCH_2_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "MOCH_2"),]
MOCH_2_tb_occur <- MOCH_2_tb[,1:91]
MOCH_2_tb_occur_len<-length(MOCH_2_tb_occur[,colSums(MOCH_2_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="MOCH_2",samples_per_class=MOCH_2_tb_occur_len))

MOCH_5_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "MOCH_5"),]
MOCH_5_tb_occur <- MOCH_5_tb[,1:91]
MOCH_5_tb_occur_len<-length(MOCH_5_tb_occur[,colSums(MOCH_5_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="MOCH_5",samples_per_class=MOCH_5_tb_occur_len))

Prasinophyceae_clade_VII_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Prasinophyceae_clade-VII"),]
Prasinophyceae_clade_VII_tb_occur <- Prasinophyceae_clade_VII_tb[,1:91]
Prasinophyceae_clade_VII_tb_occur_len<-length(Prasinophyceae_clade_VII_tb_occur[,colSums(Prasinophyceae_clade_VII_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Prasinophyceae_clade-VII",samples_per_class=Prasinophyceae_clade_VII_tb_occur_len))

Prasinophyceae_clade_IX_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Prasinophyceae_clade-IX"),]
Prasinophyceae_clade_IX_tb_occur <- Prasinophyceae_clade_IX_tb[,1:91]
Prasinophyceae_clade_IX_tb_occur_len<-length(Prasinophyceae_clade_IX_tb_occur[,colSums(Prasinophyceae_clade_IX_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Prasinophyceae_clade-IX",samples_per_class=Prasinophyceae_clade_IX_tb_occur_len))

Pyramimonadaceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Pyramimonadaceae"),]
Pyramimonadaceae_tb_occur <- Pyramimonadaceae_tb[,1:91]
Pyramimonadaceae_tb_occur_len<-length(Pyramimonadaceae_tb_occur[,colSums(Pyramimonadaceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Pyramimonadaceae",samples_per_class=Pyramimonadaceae_tb_occur_len))

other_plastids_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "other_plastids"),]
other_plastids_tb_occur <- other_plastids_tb[,1:91]
other_plastids_tb_occur_len<-length(other_plastids_tb_occur[,colSums(other_plastids_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="other_plastids",samples_per_class=other_plastids_tb_occur_len))

Cyanobacteria_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Cyanobacteria"),]
Cyanobacteria_tb_occur <- Cyanobacteria_tb[,1:91]
Cyanobacteria_tb_occur_len<-length(Cyanobacteria_tb_occur[,colSums(Cyanobacteria_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Cyanobacteria",samples_per_class=Cyanobacteria_tb_occur_len))

occurrence_counts_phototrophs
#row.names(occurrence_counts_phototrophs)<-occurrence_counts_phototrophs$group
occurrence_counts_phototrophs<-as.data.frame(occurrence_counts_phototrophs)
```


**Absolute values**

```{r aggregate_16S, echo=FALSE, results="hide", warning=FALSE}
class_summary_reads_per_class<-aggregate(rowSums(tb16_phototrophs[1:91]), list(tb16_phototrophs$class_A), sum)
# count the different groups

class_summary_otus_per_class<-aggregate(rowSums(tb16_phototrophs[1:91]), list(tb16_phototrophs$class_A), length)

## READS PER CLASS ##
attach(class_summary_reads_per_class)
class_summary_reads_per_class_order<-class_summary_reads_per_class[order(-x),]
detach(class_summary_reads_per_class)
class_summary_reads_per_class_order

#fix column names
row.names(class_summary_reads_per_class_order)<-class_summary_reads_per_class_order[,1]
class_summary_reads_per_class_order<-class_summary_reads_per_class_order[c(-1)]
colnames(class_summary_reads_per_class_order)<-c("reads_per_class")

## OTUs PER CLASS ##
attach(class_summary_otus_per_class)
class_summary_otus_per_class_order<-class_summary_otus_per_class[order(-x),]
detach(class_summary_otus_per_class)
class_summary_otus_per_class_order

row.names(class_summary_otus_per_class_order)<-class_summary_otus_per_class_order[,1]
class_summary_otus_per_class_order<-class_summary_otus_per_class_order[c(-1)]
colnames(class_summary_otus_per_class_order)<-c("OTUs_per_class")

#class_summary_reads<-aggregate(sum~class, data=otutab_full_wTax, FUN="sum") 
# sum reads different groups
```

```{r create_table_merging_#OTUs_#reads_#samples_16S, echo=FALSE, warning=F}
tb16S_OTUs_reads_samples <- merge(class_summary_reads_per_class_order, class_summary_otus_per_class_order, by="row.names")

row.names(tb16S_OTUs_reads_samples)<-tb16S_OTUs_reads_samples[,1]

tb16S_OTUs_reads_samples<-tb16S_OTUs_reads_samples[,-1]
colnames(tb16S_OTUs_reads_samples)<-c("reads_per_class","OTUs_per_class")
#tb16S_OTUs_reads_samples[1:5,1:2]

tb16S_OTUs_reads_samples<-tb16S_OTUs_reads_samples[order(tb16S_OTUs_reads_samples$reads_per_class, tb16S_OTUs_reads_samples$OTUs_per_class, decreasing = T), c(1,2)]

#add no. of samples per class
tb16S_OTUs_reads_samples<-merge(tb16S_OTUs_reads_samples, occurrence_counts_phototrophs, by.x="row.names", by.y="group", all.x=TRUE)

row.names(tb16S_OTUs_reads_samples)<-tb16S_OTUs_reads_samples[,1]
tb16S_OTUs_reads_samples<-tb16S_OTUs_reads_samples[,-1]
tb16S_OTUs_reads_samples
```


\
**Relative values**

```{r OTUs_vs_reads_relative_values_16S, echo=FALSE, warning=F}

tb16S_OTUs_reads_samples_rel_abund <- tb16S_OTUs_reads_samples

tb16S_OTUs_reads_samples_rel_abund$reads_per_class<-(tb16S_OTUs_reads_samples_rel_abund$reads_per_class*100)/colSums(tb16S_OTUs_reads_samples)[1]
tb16S_OTUs_reads_samples_rel_abund$OTUs_per_class<-(tb16S_OTUs_reads_samples_rel_abund$OTUs_per_class*100)/colSums(tb16S_OTUs_reads_samples)[2]
tb16S_OTUs_reads_samples_rel_abund$samples_per_class<-(tb16S_OTUs_reads_samples_rel_abund$samples_per_class*100)/91

colSums(tb16S_OTUs_reads_samples_rel_abund)
tb16S_OTUs_reads_samples_rel_abund
```

\
\

Reads per class vs. OTUs per class:

```{r OTUs_vs_reads_16S_v2, echo=FALSE, warnings=F, warning=F}
ggplot(tb16S_OTUs_reads_samples_rel_abund) +
  geom_point(aes(OTUs_per_class, reads_per_class)) +
  geom_text_repel(aes(OTUs_per_class, reads_per_class, label = rownames(tb16S_OTUs_reads_samples)), size=4, force = 3, box.padding = unit(0.1, "lines"), point.padding = unit(0.02, 'lines')) + labs(title="[B3] MP 16S: reads per class vs. OTUs per class",  x="\ndiversity (no. of OTUs in %)", y="abundance (no. of reads in %)\n") + scale_x_log10(limits = c(-1,200), breaks = c(0,0.1,1,10,100)) + scale_y_log10(limits = c(-1,200), breaks = c(0,0.1,1,10,100))  + theme(axis.text=element_text(size=14), axis.title=element_text(size=16))
```

\
\

Reads per class vs. samples in which they occurr:

```{r reads_vs_samples_16S, echo=FALSE, warning=F}
ggplot(tb16S_OTUs_reads_samples_rel_abund) +
  geom_point(aes(samples_per_class,reads_per_class)) +
  geom_text_repel(aes(samples_per_class,reads_per_class, label = rownames(tb16S_OTUs_reads_samples)), size=4, force = 8, box.padding = unit(0.1, "lines"), point.padding = unit(0.2, 'lines')) + scale_x_continuous(limits = c(0, 116), breaks=c(25,50,75,100)) + labs(title="[B4] MP 16S: reads per class vs. samples in which they occur", x="\nsamples in which they occurr (%)", y="abundance (no. of reads in %)\n") + scale_y_log10(limits = c(-1,200), breaks = c(0,0.1,1,10,100)) + theme(axis.text=element_text(size=14), axis.title=element_text(size=16))
```


```{r cyanobacteria_histogram, echo=F, results="hide", warning=F}
cyano_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Cyanobacteria"),]

class_summary_reads_per_class_cyano<-aggregate(rowSums(cyano_tb[1:91]), list(cyano_tb$class_B), sum)
# count the different groups
class_summary_otus_per_class_cyano<-aggregate(rowSums(cyano_tb[1:91]), list(cyano_tb$class_B), length)

## READS PER CLASS ##
attach(class_summary_reads_per_class_cyano)
class_summary_reads_per_class_cyano_order<-class_summary_reads_per_class_cyano[order(-x),]
detach(class_summary_reads_per_class_cyano)
class_summary_reads_per_class_cyano_order

#fix column names
row.names(class_summary_reads_per_class_cyano_order)<-class_summary_reads_per_class_cyano_order[,1]
class_summary_reads_per_class_cyano_order<-class_summary_reads_per_class_cyano_order[c(-1)]
colnames(class_summary_reads_per_class_cyano_order)<-c("reads_per_class")


## OTUs PER CLASS ##
attach(class_summary_otus_per_class_cyano)
class_summary_otus_per_class_cyano_order<-class_summary_otus_per_class_cyano[order(-x),]
detach(class_summary_otus_per_class_cyano)
class_summary_otus_per_class_cyano_order

row.names(class_summary_otus_per_class_cyano_order)<-class_summary_otus_per_class_cyano_order[,1]
class_summary_otus_per_class_cyano_order<-class_summary_otus_per_class_cyano_order[c(-1)]
colnames(class_summary_otus_per_class_cyano_order)<-c("OTUs_per_class")

# create_table_merging_#OTUs_#reads
cyano_OTUs_reads <- merge(class_summary_reads_per_class_cyano_order, class_summary_otus_per_class_cyano_order, by="row.names")

row.names(cyano_OTUs_reads)<-cyano_OTUs_reads[,1]
cyano_OTUs_reads<-cyano_OTUs_reads[,-1]
colnames(cyano_OTUs_reads)<-c("reads_per_class","OTUs_per_class")
cyano_OTUs_reads[1:3,1:2]

cyano_OTUs_reads<-cyano_OTUs_reads[order(cyano_OTUs_reads$reads_per_class, cyano_OTUs_reads$OTUs_per_class, decreasing = T), c(1,2)]
```

\
Absolute values of cyanobacteria groups richness and abundance:

```{r cyanos_abs_values, echo=F}
cyano_OTUs_reads
```

```{r cyanos_rel_values, echo=F, results="hide"}
#compute relative values
cyano_OTUs_reads_rel_abund <- cyano_OTUs_reads

cyano_OTUs_reads_rel_abund$reads_per_class<-(cyano_OTUs_reads_rel_abund$reads_per_class*100)/colSums(cyano_OTUs_reads)[1]
cyano_OTUs_reads_rel_abund$OTUs_per_class<-(cyano_OTUs_reads_rel_abund$OTUs_per_class*100)/colSums(cyano_OTUs_reads)[2]

colSums(cyano_OTUs_reads_rel_abund)
rownames(cyano_OTUs_reads_rel_abund) = c("Prochlorococcus", "Synechococcus", "Other cyanobacteria")
cyano_OTUs_reads_rel_abund

cyano_OTUs_reads_rel_abund["cyano_group"]<-NA
cyano_OTUs_reads_rel_abund$cyano_group<-c("Prochlorococcus", "Synechococcus", "Other cyanobacteria")

cyano_OTUs_reads_rel_abund2<-read.table("input_99/cyano_histograms_data.txt", head=TRUE)
```


Relative values of Cyanobacteria groups richness and abundance:

```{r print_rel_values, echo=F}
cyano_OTUs_reads_rel_abund[,1:2]

ggplot(cyano_OTUs_reads_rel_abund2, aes(x=group, y=value, fill=data)) + 
  geom_bar(position="dodge", stat="identity") +
  scale_y_continuous(limits=c(-1,100), breaks=c(25,50,75,100)) +
  scale_fill_manual(values=c("#000000", "cadetblue3")) +
  labs(x="", y="% \n") + 
  theme(axis.text=element_text(size=14), axis.title=element_text(size=16)) + 
  theme(legend.title = element_blank(), legend.position=c(0.85,0.85), legend.text=element_text(size=14)) +
  theme(axis.title.y = element_text(size = rel(1.5), angle = 0, vjust=0.97))
```


\
**PROTISTS**

```{r dim_merge2_16S.protists, echo=F}
#selection of phototrophs:
tb16_phototrophs <- tb16_phototrophs[which(tb16_phototrophs$class_A != "Cyanobacteria"),]

nrow(tb16_phototrophs)
sum(colSums(tb16_phototrophs[,1:91]))

#dim(tb16_phototrophs)
#tb16_phototrophs[1:5,1:5]
```

```{r count_samples_per_group_16S.protists, echo=F, warning=FALSE, results="hide"}
occurrence_counts_phototrophs<-data.table()
nrow(tb16_phototrophs)

#create a table per group and count in how many samples they occur. 
Dinophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Dinophyceae"),]
Dinophyceae_tb_occur <- Dinophyceae_tb[,1:91]
Dinophyceae_tb_occur_len<-length(Dinophyceae_tb_occur[,colSums(Dinophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Dinophyceae",samples_per_class=Dinophyceae_tb_occur_len))

Prasinophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "other_Prasinophyceae"),]
Prasinophyceae_tb_occur <- Prasinophyceae_tb[,1:91]
Prasinophyceae_tb_occur_len<-length(Prasinophyceae_tb_occur[,colSums(Prasinophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="other_Prasinophyceae",samples_per_class=Prasinophyceae_tb_occur_len))

Chrysophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Chrysophyceae"),]
Chrysophyceae_tb_occur <- Chrysophyceae_tb[,1:91]
Chrysophyceae_tb_occur_len<-length(Chrysophyceae_tb_occur[,colSums(Chrysophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Chrysophyceae",samples_per_class=Chrysophyceae_tb_occur_len))

Pelagophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Pelagophyceae"),]
Pelagophyceae_tb_occur <- Pelagophyceae_tb[,1:91]
Pelagophyceae_tb_occur_len<-length(Pelagophyceae_tb_occur[,colSums(Pelagophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Pelagophyceae",samples_per_class=Pelagophyceae_tb_occur_len))

Dictyochophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Dictyochophyceae"),]
Dictyochophyceae_tb_occur <- Dictyochophyceae_tb[,1:91]
Dictyochophyceae_tb_occur_len<-length(Dictyochophyceae_tb_occur[,colSums(Dictyochophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Dictyochophyceae",samples_per_class=Dictyochophyceae_tb_occur_len))

Cryptophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Cryptophyceae"),]
Cryptophyceae_tb_occur <- Cryptophyceae_tb[,1:91]
Cryptophyceae_tb_occur_len<-length(Cryptophyceae_tb_occur[,colSums(Cryptophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Cryptophyceae",samples_per_class=Cryptophyceae_tb_occur_len))

Bacillariophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Bacillariophyceae"),]
Bacillariophyceae_tb_occur <- Bacillariophyceae_tb[,1:91]
Bacillariophyceae_tb_occur_len<-length(Bacillariophyceae_tb_occur[,colSums(Bacillariophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Bacillariophyceae",samples_per_class=Bacillariophyceae_tb_occur_len))

Chlorarachniophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Chlorarachniophyceae"),]
Chlorarachniophyceae_tb_occur <- Chlorarachniophyceae_tb[,1:91]
Chlorarachniophyceae_tb_occur_len<-length(Chlorarachniophyceae_tb_occur[,colSums(Chlorarachniophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Chlorarachniophyceae",samples_per_class=Chlorarachniophyceae_tb_occur_len))

Bolidophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Bolidophyceae"),]
Bolidophyceae_tb_occur <- Bolidophyceae_tb[,1:91]
Bolidophyceae_tb_occur_len<-length(Bolidophyceae_tb_occur[,colSums(Bolidophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Bolidophyceae",samples_per_class=Bolidophyceae_tb_occur_len))

Pinguiophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Pinguiophyceae"),]
Pinguiophyceae_tb_occur <- Pinguiophyceae_tb[,1:91]
Pinguiophyceae_tb_occur_len<-length(Pinguiophyceae_tb_occur[,colSums(Pinguiophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Pinguiophyceae",samples_per_class=Pinguiophyceae_tb_occur_len))

Prymnesiophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Prymnesiophyceae"),]
Prymnesiophyceae_tb_occur <- Prymnesiophyceae_tb[,1:91]
Prymnesiophyceae_tb_occur_len<-length(Prymnesiophyceae_tb_occur[,colSums(Prymnesiophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Prymnesiophyceae",samples_per_class=Prymnesiophyceae_tb_occur_len))

Mamiellophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Mamiellophyceae"),]
Mamiellophyceae_tb_occur <- Mamiellophyceae_tb[,1:91]
Mamiellophyceae_tb_occur_len<-length(Mamiellophyceae_tb_occur[,colSums(Mamiellophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Mamiellophyceae",samples_per_class=Mamiellophyceae_tb_occur_len))

Eustigmatophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Eustigmatophyceae"),]
Eustigmatophyceae_tb_occur <- Eustigmatophyceae_tb[,1:91]
Eustigmatophyceae_tb_occur_len<-length(Eustigmatophyceae_tb_occur[,colSums(Eustigmatophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Eustigmatophyceae",samples_per_class=Eustigmatophyceae_tb_occur_len))

Chlorophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Chlorophyceae"),]
Chlorophyceae_tb_occur <- Chlorophyceae_tb[,1:91]
Chlorophyceae_tb_occur_len<-length(Chlorophyceae_tb_occur[,colSums(Chlorophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Chlorophyceae",samples_per_class=Chlorophyceae_tb_occur_len))

Ulvophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Ulvophyceae"),]
Ulvophyceae_tb_occur <- Ulvophyceae_tb[,1:91]
Ulvophyceae_tb_occur_len<-length(Ulvophyceae_tb_occur[,colSums(Ulvophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Ulvophyceae",samples_per_class=Ulvophyceae_tb_occur_len))

Raphydophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Raphydophyceae"),]
Raphydophyceae_tb_occur <- Raphydophyceae_tb[,1:91]
Raphydophyceae_tb_occur_len<-length(Raphydophyceae_tb_occur[,colSums(Raphydophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Raphydophyceae",samples_per_class=Raphydophyceae_tb_occur_len))

Trebouxiophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Trebouxiophyceae"),]
Trebouxiophyceae_tb_occur <- Trebouxiophyceae_tb[,1:91]
Trebouxiophyceae_tb_occur_len<-length(Trebouxiophyceae_tb_occur[,colSums(Trebouxiophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Trebouxiophyceae",samples_per_class=Trebouxiophyceae_tb_occur_len))

Phaeophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Phaeophyceae"),]
Phaeophyceae_tb_occur <- Phaeophyceae_tb[,1:91]
Phaeophyceae_tb_occur_len<-length(Phaeophyceae_tb_occur[,colSums(Phaeophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Phaeophyceae",samples_per_class=Phaeophyceae_tb_occur_len))

Phaeothamniophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Phaeothamniophyceae"),]
Phaeothamniophyceae_tb_occur <- Phaeothamniophyceae_tb[,1:91]
Phaeothamniophyceae_tb_occur_len<-length(Phaeothamniophyceae_tb_occur[,colSums(Phaeothamniophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Phaeothamniophyceae",samples_per_class=Phaeothamniophyceae_tb_occur_len))

Xanthophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Xanthophyceae"),]
Xanthophyceae_tb_occur <- Xanthophyceae_tb[,1:91]
Xanthophyceae_tb_occur_len<-length(Xanthophyceae_tb_occur[,colSums(Xanthophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Xanthophyceae",samples_per_class=Xanthophyceae_tb_occur_len))

Chlorodendrophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Chlorodendrophyceae"),]
Chlorodendrophyceae_tb_occur <- Chlorodendrophyceae_tb[,1:91]
Chlorodendrophyceae_tb_occur_len<-length(Chlorodendrophyceae_tb_occur[,colSums(Chlorodendrophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Chlorodendrophyceae",samples_per_class=Chlorodendrophyceae_tb_occur_len))

IncertaeSedis_Archaeplastida_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "IncertaeSedis_Archaeplastida"),]
IncertaeSedis_Archaeplastida_tb_occur <- IncertaeSedis_Archaeplastida_tb[,1:91]
IncertaeSedis_Archaeplastida_tb_occur_len<-length(IncertaeSedis_Archaeplastida_tb_occur[,colSums(IncertaeSedis_Archaeplastida_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="IncertaeSedis_Archaeplastida",samples_per_class=IncertaeSedis_Archaeplastida_tb_occur_len))

Nephroselmidophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Nephroselmidophyceae"),]
Nephroselmidophyceae_tb_occur <- Nephroselmidophyceae_tb[,1:91]
Nephroselmidophyceae_tb_occur_len<-length(Nephroselmidophyceae_tb_occur[,colSums(Nephroselmidophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Nephroselmidophyceae",samples_per_class=Nephroselmidophyceae_tb_occur_len))

Pavlovophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Pavlovophyceae"),]
Pavlovophyceae_tb_occur <- Pavlovophyceae_tb[,1:91]
Pavlovophyceae_tb_occur_len<-length(Pavlovophyceae_tb_occur[,colSums(Pavlovophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Pavlovophyceae",samples_per_class=Pavlovophyceae_tb_occur_len))

Rhodophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Rhodophyceae"),]
Rhodophyceae_tb_occur <- Rhodophyceae_tb[,1:91]
Rhodophyceae_tb_occur_len<-length(Rhodophyceae_tb_occur[,colSums(Rhodophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Rhodophyceae",samples_per_class=Rhodophyceae_tb_occur_len))

Rappemonads_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Rappemonads"),]
Rappemonads_tb_occur <- Rappemonads_tb[,1:91]
Rappemonads_tb_occur_len<-length(Rappemonads_tb_occur[,colSums(Rappemonads_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Rappemonads",samples_per_class=Rappemonads_tb_occur_len))

MOCH_1_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "MOCH_1"),]
MOCH_1_tb_occur <- MOCH_1_tb[,1:91]
MOCH_1_tb_occur_len<-length(MOCH_1_tb_occur[,colSums(MOCH_1_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="MOCH_1",samples_per_class=MOCH_1_tb_occur_len))

MOCH_2_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "MOCH_2"),]
MOCH_2_tb_occur <- MOCH_2_tb[,1:91]
MOCH_2_tb_occur_len<-length(MOCH_2_tb_occur[,colSums(MOCH_2_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="MOCH_2",samples_per_class=MOCH_2_tb_occur_len))

MOCH_5_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "MOCH_5"),]
MOCH_5_tb_occur <- MOCH_5_tb[,1:91]
MOCH_5_tb_occur_len<-length(MOCH_5_tb_occur[,colSums(MOCH_5_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="MOCH_5",samples_per_class=MOCH_5_tb_occur_len))

Prasinophyceae_clade_VII_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Prasinophyceae_clade-VII"),]
Prasinophyceae_clade_VII_tb_occur <- Prasinophyceae_clade_VII_tb[,1:91]
Prasinophyceae_clade_VII_tb_occur_len<-length(Prasinophyceae_clade_VII_tb_occur[,colSums(Prasinophyceae_clade_VII_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Prasinophyceae_clade-VII",samples_per_class=Prasinophyceae_clade_VII_tb_occur_len))

Prasinophyceae_clade_IX_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Prasinophyceae_clade-IX"),]
Prasinophyceae_clade_IX_tb_occur <- Prasinophyceae_clade_IX_tb[,1:91]
Prasinophyceae_clade_IX_tb_occur_len<-length(Prasinophyceae_clade_IX_tb_occur[,colSums(Prasinophyceae_clade_IX_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Prasinophyceae_clade-IX",samples_per_class=Prasinophyceae_clade_IX_tb_occur_len))

Pyramimonadaceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Pyramimonadaceae"),]
Pyramimonadaceae_tb_occur <- Pyramimonadaceae_tb[,1:91]
Pyramimonadaceae_tb_occur_len<-length(Pyramimonadaceae_tb_occur[,colSums(Pyramimonadaceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Pyramimonadaceae",samples_per_class=Pyramimonadaceae_tb_occur_len))

other_plastids_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "other_plastids"),]
other_plastids_tb_occur <- other_plastids_tb[,1:91]
other_plastids_tb_occur_len<-length(other_plastids_tb_occur[,colSums(other_plastids_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="other_plastids",samples_per_class=other_plastids_tb_occur_len))

occurrence_counts_phototrophs
#row.names(occurrence_counts_phototrophs)<-occurrence_counts_phototrophs$group
occurrence_counts_phototrophs<-as.data.frame(occurrence_counts_phototrophs)
```


**Absolute values**

```{r aggregate_16S.protists, echo=FALSE, results="hide", warning=FALSE}
class_summary_reads_per_class<-aggregate(rowSums(tb16_phototrophs[1:91]), list(tb16_phototrophs$class_A), sum)
# count the different groups

class_summary_otus_per_class<-aggregate(rowSums(tb16_phototrophs[1:91]), list(tb16_phototrophs$class_A), length)

## READS PER CLASS ##
attach(class_summary_reads_per_class)
class_summary_reads_per_class_order<-class_summary_reads_per_class[order(-x),]
detach(class_summary_reads_per_class)
class_summary_reads_per_class_order

#fix column names
row.names(class_summary_reads_per_class_order)<-class_summary_reads_per_class_order[,1]
class_summary_reads_per_class_order<-class_summary_reads_per_class_order[c(-1)]
colnames(class_summary_reads_per_class_order)<-c("reads_per_class")

## OTUs PER CLASS ##
attach(class_summary_otus_per_class)
class_summary_otus_per_class_order<-class_summary_otus_per_class[order(-x),]
detach(class_summary_otus_per_class)
class_summary_otus_per_class_order

row.names(class_summary_otus_per_class_order)<-class_summary_otus_per_class_order[,1]
class_summary_otus_per_class_order<-class_summary_otus_per_class_order[c(-1)]
colnames(class_summary_otus_per_class_order)<-c("OTUs_per_class")

#class_summary_reads<-aggregate(sum~class, data=otutab_full_wTax, FUN="sum") 
# sum reads different groups
```

```{r create_table_merging_#OTUs_#reads_#samples_16S.protists, echo=FALSE, warning=F}
tb16S_OTUs_reads_samples <- merge(class_summary_reads_per_class_order, class_summary_otus_per_class_order, by="row.names")

row.names(tb16S_OTUs_reads_samples)<-tb16S_OTUs_reads_samples[,1]

tb16S_OTUs_reads_samples<-tb16S_OTUs_reads_samples[,-1]
colnames(tb16S_OTUs_reads_samples)<-c("reads_per_class","OTUs_per_class")
#tb16S_OTUs_reads_samples[1:5,1:2]

tb16S_OTUs_reads_samples<-tb16S_OTUs_reads_samples[order(tb16S_OTUs_reads_samples$reads_per_class, tb16S_OTUs_reads_samples$OTUs_per_class, decreasing = T), c(1,2)]

#add no. of samples per class
tb16S_OTUs_reads_samples<-merge(tb16S_OTUs_reads_samples, occurrence_counts_phototrophs, by.x="row.names", by.y="group", all.x=TRUE)

row.names(tb16S_OTUs_reads_samples)<-tb16S_OTUs_reads_samples[,1]
tb16S_OTUs_reads_samples<-tb16S_OTUs_reads_samples[,-1]
tb16S_OTUs_reads_samples
```


\
**Relative values**

```{r OTUs_vs_reads_relative_values_16S.protists, echo=FALSE, warning=F}

tb16S_OTUs_reads_samples_rel_abund <- tb16S_OTUs_reads_samples

tb16S_OTUs_reads_samples_rel_abund$reads_per_class<-(tb16S_OTUs_reads_samples_rel_abund$reads_per_class*100)/colSums(tb16S_OTUs_reads_samples)[1]
tb16S_OTUs_reads_samples_rel_abund$OTUs_per_class<-(tb16S_OTUs_reads_samples_rel_abund$OTUs_per_class*100)/colSums(tb16S_OTUs_reads_samples)[2]
tb16S_OTUs_reads_samples_rel_abund$samples_per_class<-(tb16S_OTUs_reads_samples_rel_abund$samples_per_class*100)/91

colSums(tb16S_OTUs_reads_samples_rel_abund)
tb16S_OTUs_reads_samples_rel_abund
```

\
\

Reads per class vs. OTUs per class:

```{r OTUs_vs_reads_16S_v2.protists, echo=FALSE, warnings=F, warning=F}
ggplot(tb16S_OTUs_reads_samples_rel_abund) +
  geom_point(aes(OTUs_per_class, reads_per_class)) +
  geom_text_repel(aes(OTUs_per_class, reads_per_class, label = rownames(tb16S_OTUs_reads_samples)), size=4, force = 3, box.padding = unit(0.1, "lines"), point.padding = unit(0.02, 'lines')) + labs(title="[B3.2] MP 16S: reads per class vs. OTUs per class",  x="\ndiversity (no. of OTUs in %)", y="abundance (no. of reads in %)\n") + scale_x_log10(limits = c(-1,200), breaks = c(0,0.1,1,10,100)) + scale_y_log10(limits = c(-1,200), breaks = c(0,0.1,1,10,100))  + theme(axis.text=element_text(size=14), axis.title=element_text(size=16))
```

\
\

Reads per class vs. samples in which they occurr:

```{r reads_vs_samples_16S.protists, echo=FALSE, warning=F}
ggplot(tb16S_OTUs_reads_samples_rel_abund) +
  geom_point(aes(samples_per_class,reads_per_class)) +
  geom_text_repel(aes(samples_per_class,reads_per_class, label = rownames(tb16S_OTUs_reads_samples)), size=4, force = 8, box.padding = unit(0.1, "lines"), point.padding = unit(0.2, 'lines')) + scale_x_continuous(limits = c(0, 116), breaks=c(25,50,75,100)) + labs(title="[B4.2] MP 16S: reads per class vs. OTUs per class",  x="\ndiversity (no. of OTUs in %)", y="abundance (no. of reads in %)\n") + scale_y_log10(limits = c(-1,200), breaks = c(0,0.1,1,10,100)) + theme(axis.text=element_text(size=14), axis.title=element_text(size=16))
```
\




### 2.6.2) Non-rarefied data


```{r merge_tables_non.raref_16S, echo=FALSE, results="hide", warning=FALSE}
#Let's add the taxonomic class_Aication by merging "tb16_tax_occur_min16509_no_cero" with "tb16_tax":

tb16_tax_occur_min16509_t<-tb16_tax_occur_min16509
tb16_min16509_tax<-merge(tb16_tax_occur_min16509_t,tb16_class, by="row.names")


dim(tb16_min16509_tax)
tb16_min16509_tax[1:5,1:5]

#fix OTU_no as new row
rownames(tb16_min16509_tax)=tb16_min16509_tax$Row.names
#add OTU_no as rowname
rownames.tb16_min16509_tax<-tb16_min16509_tax[,1]
tb16_min16509_tax<-tb16_min16509_tax[,-1]

dim(tb16_min16509_tax)
tb16_min16509_tax[1:5, 1:5]

#sort by OTU_no (split rowname, introduce no. into column "OTU_no" and sort)
tb16_min16509_tax["OTU_no"] <- NA
tb16_min16509_tax$OTU_no <- sapply(strsplit(rownames(tb16_min16509_tax),split= "\\_"),'[',2)
tb16_min16509_tax$OTU_no <- as.numeric(as.character(tb16_min16509_tax$OTU_no))
tb16_min16509_tax_sorted<-tb16_min16509_tax[order(tb16_min16509_tax$OTU_no, decreasing = FALSE), ]
```

No. of OTUs and reads of the rearefied dataset:

```{r dim_merge1_non.raref_16S, echo=F}
nrow(tb16_min16509_tax_sorted)
sum(colSums(tb16_min16509_tax_sorted[,1:91]))

#dim(tb16_min16509_tax_sorted)
#tb16_min16509_tax_sorted[1:5,1:5]
```

\
Number and proportion (%) of OTUs of phototrophic groups:

```{r dim_merge2_non.raref, echo=F}
#selection of phototrophs:
tb16_phototrophs <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A != "other_bacteria"),]

nrow(tb16_phototrophs)
nrow(tb16_phototrophs)*100/nrow(tb16_min16509_tax_sorted)
```

Number and proportion (%) of reads of phototrophic groups:

```{r dim_merge2.2_non.raref, echo=F}
sum(colSums(tb16_phototrophs[,1:43]))
(sum(colSums(tb16_phototrophs[,1:43])))*100/sum(colSums(tb16_min16509_tax_sorted[,1:43]))

#dim(tb16_phototrophs)
#tb16_phototrophs[1:5,1:5]
```

\
Number and proportion (%) of OTUs of non-phototrophic groups:

```{r dim_merge3_non.raref, echo=F}
#selection of non.phototrophs:
tb16_non.phototrophs <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "other_bacteria"),]

nrow(tb16_non.phototrophs)
(nrow(tb16_non.phototrophs))*100/nrow(tb16_min16509_tax_sorted)
```

Number and proportion (%) of reads of non-phototrophic groups:

```{r dim_merge3.2_non.raref, echo=F}
sum(colSums(tb16_non.phototrophs[,1:43]))
(sum(colSums(tb16_non.phototrophs[,1:43])))*100/sum(colSums(tb16_min16509_tax_sorted[,1:43]))
```


\
**PHOTOTROPHS + HETEROTROPHS**

**Absolute values**


```{r count_samples_per_group_Heterotrophs_non.raref_16S, echo=F, warning=FALSE, results="hide"}
occurrence_counts_phototrophs<-data.table()

#create a table per group and count in how many samples they occur. 
Dinophyceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Dinophyceae"),]
Dinophyceae_tb_occur <- Dinophyceae_tb[,1:91]
Dinophyceae_tb_occur_len<-length(Dinophyceae_tb_occur[,colSums(Dinophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Dinophyceae",samples_per_class=Dinophyceae_tb_occur_len))

Prasinophyceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "other_Prasinophyceae"),]
Prasinophyceae_tb_occur <- Prasinophyceae_tb[,1:91]
Prasinophyceae_tb_occur_len<-length(Prasinophyceae_tb_occur[,colSums(Prasinophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="other_Prasinophyceae",samples_per_class=Prasinophyceae_tb_occur_len))

Chrysophyceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Chrysophyceae"),]
Chrysophyceae_tb_occur <- Chrysophyceae_tb[,1:91]
Chrysophyceae_tb_occur_len<-length(Chrysophyceae_tb_occur[,colSums(Chrysophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Chrysophyceae",samples_per_class=Chrysophyceae_tb_occur_len))

Pelagophyceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Pelagophyceae"),]
Pelagophyceae_tb_occur <- Pelagophyceae_tb[,1:91]
Pelagophyceae_tb_occur_len<-length(Pelagophyceae_tb_occur[,colSums(Pelagophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Pelagophyceae",samples_per_class=Pelagophyceae_tb_occur_len))

Dictyochophyceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Dictyochophyceae"),]
Dictyochophyceae_tb_occur <- Dictyochophyceae_tb[,1:91]
Dictyochophyceae_tb_occur_len<-length(Dictyochophyceae_tb_occur[,colSums(Dictyochophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Dictyochophyceae",samples_per_class=Dictyochophyceae_tb_occur_len))

Cryptophyceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Cryptophyceae"),]
Cryptophyceae_tb_occur <- Cryptophyceae_tb[,1:91]
Cryptophyceae_tb_occur_len<-length(Cryptophyceae_tb_occur[,colSums(Cryptophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Cryptophyceae",samples_per_class=Cryptophyceae_tb_occur_len))

Bacillariophyceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Bacillariophyceae"),]
Bacillariophyceae_tb_occur <- Bacillariophyceae_tb[,1:91]
Bacillariophyceae_tb_occur_len<-length(Bacillariophyceae_tb_occur[,colSums(Bacillariophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Bacillariophyceae",samples_per_class=Bacillariophyceae_tb_occur_len))

Chlorarachniophyceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Chlorarachniophyceae"),]
Chlorarachniophyceae_tb_occur <- Chlorarachniophyceae_tb[,1:91]
Chlorarachniophyceae_tb_occur_len<-length(Chlorarachniophyceae_tb_occur[,colSums(Chlorarachniophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Chlorarachniophyceae",samples_per_class=Chlorarachniophyceae_tb_occur_len))

Bolidophyceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Bolidophyceae"),]
Bolidophyceae_tb_occur <- Bolidophyceae_tb[,1:91]
Bolidophyceae_tb_occur_len<-length(Bolidophyceae_tb_occur[,colSums(Bolidophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Bolidophyceae",samples_per_class=Bolidophyceae_tb_occur_len))

Pinguiophyceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Pinguiophyceae"),]
Pinguiophyceae_tb_occur <- Pinguiophyceae_tb[,1:91]
Pinguiophyceae_tb_occur_len<-length(Pinguiophyceae_tb_occur[,colSums(Pinguiophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Pinguiophyceae",samples_per_class=Pinguiophyceae_tb_occur_len))

Prymnesiophyceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Prymnesiophyceae"),]
Prymnesiophyceae_tb_occur <- Prymnesiophyceae_tb[,1:91]
Prymnesiophyceae_tb_occur_len<-length(Prymnesiophyceae_tb_occur[,colSums(Prymnesiophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Prymnesiophyceae",samples_per_class=Prymnesiophyceae_tb_occur_len))

Mamiellophyceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Mamiellophyceae"),]
Mamiellophyceae_tb_occur <- Mamiellophyceae_tb[,1:91]
Mamiellophyceae_tb_occur_len<-length(Mamiellophyceae_tb_occur[,colSums(Mamiellophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Mamiellophyceae",samples_per_class=Mamiellophyceae_tb_occur_len))

Eustigmatophyceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Eustigmatophyceae"),]
Eustigmatophyceae_tb_occur <- Eustigmatophyceae_tb[,1:91]
Eustigmatophyceae_tb_occur_len<-length(Eustigmatophyceae_tb_occur[,colSums(Eustigmatophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Eustigmatophyceae",samples_per_class=Eustigmatophyceae_tb_occur_len))

Chlorophyceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Chlorophyceae"),]
Chlorophyceae_tb_occur <- Chlorophyceae_tb[,1:91]
Chlorophyceae_tb_occur_len<-length(Chlorophyceae_tb_occur[,colSums(Chlorophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Chlorophyceae",samples_per_class=Chlorophyceae_tb_occur_len))

Ulvophyceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Ulvophyceae"),]
Ulvophyceae_tb_occur <- Ulvophyceae_tb[,1:91]
Ulvophyceae_tb_occur_len<-length(Ulvophyceae_tb_occur[,colSums(Ulvophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Ulvophyceae",samples_per_class=Ulvophyceae_tb_occur_len))

Raphydophyceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Raphydophyceae"),]
Raphydophyceae_tb_occur <- Raphydophyceae_tb[,1:91]
Raphydophyceae_tb_occur_len<-length(Raphydophyceae_tb_occur[,colSums(Raphydophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Raphydophyceae",samples_per_class=Raphydophyceae_tb_occur_len))

Trebouxiophyceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Trebouxiophyceae"),]
Trebouxiophyceae_tb_occur <- Trebouxiophyceae_tb[,1:91]
Trebouxiophyceae_tb_occur_len<-length(Trebouxiophyceae_tb_occur[,colSums(Trebouxiophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Trebouxiophyceae",samples_per_class=Trebouxiophyceae_tb_occur_len))

Phaeophyceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Phaeophyceae"),]
Phaeophyceae_tb_occur <- Phaeophyceae_tb[,1:91]
Phaeophyceae_tb_occur_len<-length(Phaeophyceae_tb_occur[,colSums(Phaeophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Phaeophyceae",samples_per_class=Phaeophyceae_tb_occur_len))

Phaeothamniophyceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Phaeothamniophyceae"),]
Phaeothamniophyceae_tb_occur <- Phaeothamniophyceae_tb[,1:91]
Phaeothamniophyceae_tb_occur_len<-length(Phaeothamniophyceae_tb_occur[,colSums(Phaeothamniophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Phaeothamniophyceae",samples_per_class=Phaeothamniophyceae_tb_occur_len))

Xanthophyceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Xanthophyceae"),]
Xanthophyceae_tb_occur <- Xanthophyceae_tb[,1:91]
Xanthophyceae_tb_occur_len<-length(Xanthophyceae_tb_occur[,colSums(Xanthophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Xanthophyceae",samples_per_class=Xanthophyceae_tb_occur_len))

Chlorodendrophyceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Chlorodendrophyceae"),]
Chlorodendrophyceae_tb_occur <- Chlorodendrophyceae_tb[,1:91]
Chlorodendrophyceae_tb_occur_len<-length(Chlorodendrophyceae_tb_occur[,colSums(Chlorodendrophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Chlorodendrophyceae",samples_per_class=Chlorodendrophyceae_tb_occur_len))

IncertaeSedis_Archaeplastida_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "IncertaeSedis_Archaeplastida"),]
IncertaeSedis_Archaeplastida_tb_occur <- IncertaeSedis_Archaeplastida_tb[,1:91]
IncertaeSedis_Archaeplastida_tb_occur_len<-length(IncertaeSedis_Archaeplastida_tb_occur[,colSums(IncertaeSedis_Archaeplastida_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="IncertaeSedis_Archaeplastida",samples_per_class=IncertaeSedis_Archaeplastida_tb_occur_len))

Nephroselmidophyceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Nephroselmidophyceae"),]
Nephroselmidophyceae_tb_occur <- Nephroselmidophyceae_tb[,1:91]
Nephroselmidophyceae_tb_occur_len<-length(Nephroselmidophyceae_tb_occur[,colSums(Nephroselmidophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Nephroselmidophyceae",samples_per_class=Nephroselmidophyceae_tb_occur_len))

Pavlovophyceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Pavlovophyceae"),]
Pavlovophyceae_tb_occur <- Pavlovophyceae_tb[,1:91]
Pavlovophyceae_tb_occur_len<-length(Pavlovophyceae_tb_occur[,colSums(Pavlovophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Pavlovophyceae",samples_per_class=Pavlovophyceae_tb_occur_len))

Rhodophyceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Rhodophyceae"),]
Rhodophyceae_tb_occur <- Rhodophyceae_tb[,1:91]
Rhodophyceae_tb_occur_len<-length(Rhodophyceae_tb_occur[,colSums(Rhodophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Rhodophyceae",samples_per_class=Rhodophyceae_tb_occur_len))

Rappemonads_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Rappemonads"),]
Rappemonads_tb_occur <- Rappemonads_tb[,1:91]
Rappemonads_tb_occur_len<-length(Rappemonads_tb_occur[,colSums(Rappemonads_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Rappemonads",samples_per_class=Rappemonads_tb_occur_len))

MOCH_1_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "MOCH_1"),]
MOCH_1_tb_occur <- MOCH_1_tb[,1:91]
MOCH_1_tb_occur_len<-length(MOCH_1_tb_occur[,colSums(MOCH_1_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="MOCH_1",samples_per_class=MOCH_1_tb_occur_len))

MOCH_2_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "MOCH_2"),]
MOCH_2_tb_occur <- MOCH_2_tb[,1:91]
MOCH_2_tb_occur_len<-length(MOCH_2_tb_occur[,colSums(MOCH_2_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="MOCH_2",samples_per_class=MOCH_2_tb_occur_len))

MOCH_5_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "MOCH_5"),]
MOCH_5_tb_occur <- MOCH_5_tb[,1:91]
MOCH_5_tb_occur_len<-length(MOCH_5_tb_occur[,colSums(MOCH_5_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="MOCH_5",samples_per_class=MOCH_5_tb_occur_len))

Prasinophyceae_clade_VII_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Prasinophyceae_clade-VII"),]
Prasinophyceae_clade_VII_tb_occur <- Prasinophyceae_clade_VII_tb[,1:91]
Prasinophyceae_clade_VII_tb_occur_len<-length(Prasinophyceae_clade_VII_tb_occur[,colSums(Prasinophyceae_clade_VII_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Prasinophyceae_clade-VII",samples_per_class=Prasinophyceae_clade_VII_tb_occur_len))

Prasinophyceae_clade_IX_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Prasinophyceae_clade-IX"),]
Prasinophyceae_clade_IX_tb_occur <- Prasinophyceae_clade_IX_tb[,1:91]
Prasinophyceae_clade_IX_tb_occur_len<-length(Prasinophyceae_clade_IX_tb_occur[,colSums(Prasinophyceae_clade_IX_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Prasinophyceae_clade-IX",samples_per_class=Prasinophyceae_clade_IX_tb_occur_len))

Pyramimonadaceae_tb <- tb16_min16509_tax_sorted[which(tb16_min16509_tax_sorted$class_A == "Pyramimonadaceae"),]
Pyramimonadaceae_tb_occur <- Pyramimonadaceae_tb[,1:91]
Pyramimonadaceae_tb_occur_len<-length(Pyramimonadaceae_tb_occur[,colSums(Pyramimonadaceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Pyramimonadaceae",samples_per_class=Pyramimonadaceae_tb_occur_len))

Cyanobacteria_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "Cyanobacteria"),]
Cyanobacteria_tb_occur <- Cyanobacteria_tb[,1:91]
Cyanobacteria_tb_occur_len<-length(Cyanobacteria_tb_occur[,colSums(Cyanobacteria_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Cyanobacteria",samples_per_class=Cyanobacteria_tb_occur_len))

other_cyanobacteria_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "other_cyanobacteria"),]
other_cyanobacteria_tb_occur <- other_cyanobacteria_tb[,1:91]
other_cyanobacteria_tb_occur_len<-length(other_cyanobacteria_tb_occur[,colSums(other_cyanobacteria_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="other_cyanobacteria",samples_per_class=other_cyanobacteria_tb_occur_len))

other_bacteria_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "other_bacteria"),]
other_bacteria_tb_occur <- other_bacteria_tb[,1:91]
other_bacteria_tb_occur_len<-length(other_bacteria_tb_occur[,colSums(other_bacteria_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="other_bacteria",samples_per_class=other_bacteria_tb_occur_len))

other_plastids_tb <- tb16_ss16509_tax_sorted[which(tb16_ss16509_tax_sorted$class_A == "other_plastids"),]
other_plastids_tb_occur <- other_plastids_tb[,1:91]
other_plastids_tb_occur_len<-length(other_plastids_tb_occur[,colSums(other_plastids_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="other_plastids",samples_per_class=other_plastids_tb_occur_len))

occurrence_counts_phototrophs
#row.names(occurrence_counts_phototrophs)<-occurrence_counts_phototrophs$group
occurrence_counts_phototrophs<-as.data.frame(occurrence_counts_phototrophs)
```

```{r aggregate_Heterotrophs_non.raref_16S, echo=FALSE, results="hide", warning=FALSE}
class_summary_reads_per_class<-aggregate(rowSums(tb16_min16509_tax_sorted[1:91]), list(tb16_min16509_tax_sorted$class_A), sum)
# count the different groups

class_summary_otus_per_class<-aggregate(rowSums(tb16_min16509_tax_sorted[1:91]), list(tb16_min16509_tax_sorted$class_A), length)

## READS PER CLASS ##
attach(class_summary_reads_per_class)
class_summary_reads_per_class_order<-class_summary_reads_per_class[order(-x),]
detach(class_summary_reads_per_class)
class_summary_reads_per_class_order

#fix column names
row.names(class_summary_reads_per_class_order)<-class_summary_reads_per_class_order[,1]
class_summary_reads_per_class_order<-class_summary_reads_per_class_order[c(-1)]
colnames(class_summary_reads_per_class_order)<-c("reads_per_class")

## OTUs PER CLASS ##
attach(class_summary_otus_per_class)
class_summary_otus_per_class_order<-class_summary_otus_per_class[order(-x),]
detach(class_summary_otus_per_class)
class_summary_otus_per_class_order

row.names(class_summary_otus_per_class_order)<-class_summary_otus_per_class_order[,1]
class_summary_otus_per_class_order<-class_summary_otus_per_class_order[c(-1)]
colnames(class_summary_otus_per_class_order)<-c("OTUs_per_class")

#class_summary_reads<-aggregate(sum~class, data=otutab_full_wTax, FUN="sum") 
# sum reads different groups
```

```{r create_table_merging_#OTUs_#reads_#samples_Heterotrophs_non.raref_16S, echo=FALSE, warning=F}
tb16S_OTUs_reads_samples <- merge(class_summary_reads_per_class_order, class_summary_otus_per_class_order, by="row.names")

row.names(tb16S_OTUs_reads_samples)<-tb16S_OTUs_reads_samples[,1]

tb16S_OTUs_reads_samples<-tb16S_OTUs_reads_samples[,-1]
colnames(tb16S_OTUs_reads_samples)<-c("reads_per_class","OTUs_per_class")
#tb16S_OTUs_reads_samples[1:5,1:2]

tb16S_OTUs_reads_samples<-tb16S_OTUs_reads_samples[order(tb16S_OTUs_reads_samples$reads_per_class, tb16S_OTUs_reads_samples$OTUs_per_class, decreasing = T), c(1,2)]

#add no. of samples per class
tb16S_OTUs_reads_samples<-merge(tb16S_OTUs_reads_samples, occurrence_counts_phototrophs, by.x="row.names", by.y="group", all.x=TRUE)

row.names(tb16S_OTUs_reads_samples)<-tb16S_OTUs_reads_samples[,1]
tb16S_OTUs_reads_samples<-tb16S_OTUs_reads_samples[,-1]
tb16S_OTUs_reads_samples
```


\
**Relative values**

```{r OTUs_vs_reads_relative_values_18S_Heterotrophs_non.raref_16S, echo=FALSE, warning=F}

tb16S_OTUs_reads_samples_rel_abund <- tb16S_OTUs_reads_samples

tb16S_OTUs_reads_samples_rel_abund$reads_per_class<-(tb16S_OTUs_reads_samples_rel_abund$reads_per_class*100)/colSums(tb16S_OTUs_reads_samples)[1]
tb16S_OTUs_reads_samples_rel_abund$OTUs_per_class<-(tb16S_OTUs_reads_samples_rel_abund$OTUs_per_class*100)/colSums(tb16S_OTUs_reads_samples)[2]
tb16S_OTUs_reads_samples_rel_abund$samples_per_class<-(tb16S_OTUs_reads_samples_rel_abund$samples_per_class*100)/91

colSums(tb16S_OTUs_reads_samples_rel_abund)
tb16S_OTUs_reads_samples_rel_abund
```

\
\

Reads per class vs. OTUs per class:

```{r OTUs_vs_reads_18S_v2_Heterotrophs_non.raref_16S, echo=FALSE, warnings=F, warning=F}
ggplot(tb16S_OTUs_reads_samples_rel_abund) +
  geom_point(aes(OTUs_per_class, reads_per_class)) +
  geom_text_repel(aes(OTUs_per_class, reads_per_class, label = rownames(tb16S_OTUs_reads_samples)), size=4, force = 3, box.padding = unit(0.1, "lines"), point.padding = unit(0.02, 'lines')) + labs(title="[B5] METAG 16S: reads per class vs. OTUs per class", x="\ndiversity (no. of OTUs in %)", y="abundance (no. of reads in %)\n") + scale_x_log10(limits = c(-1,200), breaks = c(0,0.1,1,10,100)) + scale_y_log10(limits = c(-1,200), breaks = c(0,0.1,1,10,100))  + theme(axis.text=element_text(size=14), axis.title=element_text(size=16))
```

\
\

Reads per class vs. samples in which they occurr:

```{r reads_vs_samples_18S_Heterotrophs_non.raref_16S, echo=FALSE, warning=F}
ggplot(tb16S_OTUs_reads_samples_rel_abund) +
  geom_point(aes(samples_per_class,reads_per_class)) +
  geom_text_repel(aes(samples_per_class,reads_per_class, label = rownames(tb16S_OTUs_reads_samples)), size=4, force = 8, box.padding = unit(0.1, "lines"), point.padding = unit(0.2, 'lines')) + scale_x_continuous(limits = c(0, 116), breaks=c(25,50,75,100)) + labs(title="[B6] MP 16S: reads per class vs. samples in which they occur", x="\nsamples in which they occurr (%)", y="abundance (no. of reads in %)\n") + scale_y_log10(limits = c(-1,200), breaks = c(0,0.1,1,10,100)) + theme(axis.text=element_text(size=14), axis.title=element_text(size=16))
```

\
**PHOTOTROPHS**

```{r count_samples_per_group_non.raref_16S, echo=F, warning=FALSE, results="hide"}
occurrence_counts_phototrophs<-data.table()

#create a table per group and count in how many samples they occur. 
Dinophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Dinophyceae"),]
Dinophyceae_tb_occur <- Dinophyceae_tb[,1:91]
Dinophyceae_tb_occur_len<-length(Dinophyceae_tb_occur[,colSums(Dinophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Dinophyceae",samples_per_class=Dinophyceae_tb_occur_len))

Prasinophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "other_Prasinophyceae"),]
Prasinophyceae_tb_occur <- Prasinophyceae_tb[,1:91]
Prasinophyceae_tb_occur_len<-length(Prasinophyceae_tb_occur[,colSums(Prasinophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="other_Prasinophyceae",samples_per_class=Prasinophyceae_tb_occur_len))

Chrysophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Chrysophyceae"),]
Chrysophyceae_tb_occur <- Chrysophyceae_tb[,1:91]
Chrysophyceae_tb_occur_len<-length(Chrysophyceae_tb_occur[,colSums(Chrysophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Chrysophyceae",samples_per_class=Chrysophyceae_tb_occur_len))

Pelagophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Pelagophyceae"),]
Pelagophyceae_tb_occur <- Pelagophyceae_tb[,1:91]
Pelagophyceae_tb_occur_len<-length(Pelagophyceae_tb_occur[,colSums(Pelagophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Pelagophyceae",samples_per_class=Pelagophyceae_tb_occur_len))

Dictyochophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Dictyochophyceae"),]
Dictyochophyceae_tb_occur <- Dictyochophyceae_tb[,1:91]
Dictyochophyceae_tb_occur_len<-length(Dictyochophyceae_tb_occur[,colSums(Dictyochophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Dictyochophyceae",samples_per_class=Dictyochophyceae_tb_occur_len))

Cryptophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Cryptophyceae"),]
Cryptophyceae_tb_occur <- Cryptophyceae_tb[,1:91]
Cryptophyceae_tb_occur_len<-length(Cryptophyceae_tb_occur[,colSums(Cryptophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Cryptophyceae",samples_per_class=Cryptophyceae_tb_occur_len))

Bacillariophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Bacillariophyceae"),]
Bacillariophyceae_tb_occur <- Bacillariophyceae_tb[,1:91]
Bacillariophyceae_tb_occur_len<-length(Bacillariophyceae_tb_occur[,colSums(Bacillariophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Bacillariophyceae",samples_per_class=Bacillariophyceae_tb_occur_len))

Chlorarachniophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Chlorarachniophyceae"),]
Chlorarachniophyceae_tb_occur <- Chlorarachniophyceae_tb[,1:91]
Chlorarachniophyceae_tb_occur_len<-length(Chlorarachniophyceae_tb_occur[,colSums(Chlorarachniophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Chlorarachniophyceae",samples_per_class=Chlorarachniophyceae_tb_occur_len))

Bolidophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Bolidophyceae"),]
Bolidophyceae_tb_occur <- Bolidophyceae_tb[,1:91]
Bolidophyceae_tb_occur_len<-length(Bolidophyceae_tb_occur[,colSums(Bolidophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Bolidophyceae",samples_per_class=Bolidophyceae_tb_occur_len))

Pinguiophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Pinguiophyceae"),]
Pinguiophyceae_tb_occur <- Pinguiophyceae_tb[,1:91]
Pinguiophyceae_tb_occur_len<-length(Pinguiophyceae_tb_occur[,colSums(Pinguiophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Pinguiophyceae",samples_per_class=Pinguiophyceae_tb_occur_len))

Prymnesiophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Prymnesiophyceae"),]
Prymnesiophyceae_tb_occur <- Prymnesiophyceae_tb[,1:91]
Prymnesiophyceae_tb_occur_len<-length(Prymnesiophyceae_tb_occur[,colSums(Prymnesiophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Prymnesiophyceae",samples_per_class=Prymnesiophyceae_tb_occur_len))

Mamiellophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Mamiellophyceae"),]
Mamiellophyceae_tb_occur <- Mamiellophyceae_tb[,1:91]
Mamiellophyceae_tb_occur_len<-length(Mamiellophyceae_tb_occur[,colSums(Mamiellophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Mamiellophyceae",samples_per_class=Mamiellophyceae_tb_occur_len))

Eustigmatophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Eustigmatophyceae"),]
Eustigmatophyceae_tb_occur <- Eustigmatophyceae_tb[,1:91]
Eustigmatophyceae_tb_occur_len<-length(Eustigmatophyceae_tb_occur[,colSums(Eustigmatophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Eustigmatophyceae",samples_per_class=Eustigmatophyceae_tb_occur_len))

Chlorophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Chlorophyceae"),]
Chlorophyceae_tb_occur <- Chlorophyceae_tb[,1:91]
Chlorophyceae_tb_occur_len<-length(Chlorophyceae_tb_occur[,colSums(Chlorophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Chlorophyceae",samples_per_class=Chlorophyceae_tb_occur_len))

Ulvophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Ulvophyceae"),]
Ulvophyceae_tb_occur <- Ulvophyceae_tb[,1:91]
Ulvophyceae_tb_occur_len<-length(Ulvophyceae_tb_occur[,colSums(Ulvophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Ulvophyceae",samples_per_class=Ulvophyceae_tb_occur_len))

Raphydophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Raphydophyceae"),]
Raphydophyceae_tb_occur <- Raphydophyceae_tb[,1:91]
Raphydophyceae_tb_occur_len<-length(Raphydophyceae_tb_occur[,colSums(Raphydophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Raphydophyceae",samples_per_class=Raphydophyceae_tb_occur_len))

Trebouxiophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Trebouxiophyceae"),]
Trebouxiophyceae_tb_occur <- Trebouxiophyceae_tb[,1:91]
Trebouxiophyceae_tb_occur_len<-length(Trebouxiophyceae_tb_occur[,colSums(Trebouxiophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Trebouxiophyceae",samples_per_class=Trebouxiophyceae_tb_occur_len))

Phaeophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Phaeophyceae"),]
Phaeophyceae_tb_occur <- Phaeophyceae_tb[,1:91]
Phaeophyceae_tb_occur_len<-length(Phaeophyceae_tb_occur[,colSums(Phaeophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Phaeophyceae",samples_per_class=Phaeophyceae_tb_occur_len))

Phaeothamniophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Phaeothamniophyceae"),]
Phaeothamniophyceae_tb_occur <- Phaeothamniophyceae_tb[,1:91]
Phaeothamniophyceae_tb_occur_len<-length(Phaeothamniophyceae_tb_occur[,colSums(Phaeothamniophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Phaeothamniophyceae",samples_per_class=Phaeothamniophyceae_tb_occur_len))

Xanthophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Xanthophyceae"),]
Xanthophyceae_tb_occur <- Xanthophyceae_tb[,1:91]
Xanthophyceae_tb_occur_len<-length(Xanthophyceae_tb_occur[,colSums(Xanthophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Xanthophyceae",samples_per_class=Xanthophyceae_tb_occur_len))

Chlorodendrophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Chlorodendrophyceae"),]
Chlorodendrophyceae_tb_occur <- Chlorodendrophyceae_tb[,1:91]
Chlorodendrophyceae_tb_occur_len<-length(Chlorodendrophyceae_tb_occur[,colSums(Chlorodendrophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Chlorodendrophyceae",samples_per_class=Chlorodendrophyceae_tb_occur_len))

IncertaeSedis_Archaeplastida_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "IncertaeSedis_Archaeplastida"),]
IncertaeSedis_Archaeplastida_tb_occur <- IncertaeSedis_Archaeplastida_tb[,1:91]
IncertaeSedis_Archaeplastida_tb_occur_len<-length(IncertaeSedis_Archaeplastida_tb_occur[,colSums(IncertaeSedis_Archaeplastida_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="IncertaeSedis_Archaeplastida",samples_per_class=IncertaeSedis_Archaeplastida_tb_occur_len))

Nephroselmidophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Nephroselmidophyceae"),]
Nephroselmidophyceae_tb_occur <- Nephroselmidophyceae_tb[,1:91]
Nephroselmidophyceae_tb_occur_len<-length(Nephroselmidophyceae_tb_occur[,colSums(Nephroselmidophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Nephroselmidophyceae",samples_per_class=Nephroselmidophyceae_tb_occur_len))

Pavlovophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Pavlovophyceae"),]
Pavlovophyceae_tb_occur <- Pavlovophyceae_tb[,1:91]
Pavlovophyceae_tb_occur_len<-length(Pavlovophyceae_tb_occur[,colSums(Pavlovophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Pavlovophyceae",samples_per_class=Pavlovophyceae_tb_occur_len))

Rhodophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Rhodophyceae"),]
Rhodophyceae_tb_occur <- Rhodophyceae_tb[,1:91]
Rhodophyceae_tb_occur_len<-length(Rhodophyceae_tb_occur[,colSums(Rhodophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Rhodophyceae",samples_per_class=Rhodophyceae_tb_occur_len))

Rappemonads_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Rappemonads"),]
Rappemonads_tb_occur <- Rappemonads_tb[,1:91]
Rappemonads_tb_occur_len<-length(Rappemonads_tb_occur[,colSums(Rappemonads_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Rappemonads",samples_per_class=Rappemonads_tb_occur_len))

MOCH_1_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "MOCH_1"),]
MOCH_1_tb_occur <- MOCH_1_tb[,1:91]
MOCH_1_tb_occur_len<-length(MOCH_1_tb_occur[,colSums(MOCH_1_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="MOCH_1",samples_per_class=MOCH_1_tb_occur_len))

MOCH_2_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "MOCH_2"),]
MOCH_2_tb_occur <- MOCH_2_tb[,1:91]
MOCH_2_tb_occur_len<-length(MOCH_2_tb_occur[,colSums(MOCH_2_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="MOCH_2",samples_per_class=MOCH_2_tb_occur_len))

MOCH_5_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "MOCH_5"),]
MOCH_5_tb_occur <- MOCH_5_tb[,1:91]
MOCH_5_tb_occur_len<-length(MOCH_5_tb_occur[,colSums(MOCH_5_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="MOCH_5",samples_per_class=MOCH_5_tb_occur_len))

Prasinophyceae_clade_VII_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Prasinophyceae_clade-VII"),]
Prasinophyceae_clade_VII_tb_occur <- Prasinophyceae_clade_VII_tb[,1:91]
Prasinophyceae_clade_VII_tb_occur_len<-length(Prasinophyceae_clade_VII_tb_occur[,colSums(Prasinophyceae_clade_VII_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Prasinophyceae_clade-VII",samples_per_class=Prasinophyceae_clade_VII_tb_occur_len))

Prasinophyceae_clade_IX_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Prasinophyceae_clade-IX"),]
Prasinophyceae_clade_IX_tb_occur <- Prasinophyceae_clade_IX_tb[,1:91]
Prasinophyceae_clade_IX_tb_occur_len<-length(Prasinophyceae_clade_IX_tb_occur[,colSums(Prasinophyceae_clade_IX_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Prasinophyceae_clade-IX",samples_per_class=Prasinophyceae_clade_IX_tb_occur_len))

Pyramimonadaceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Pyramimonadaceae"),]
Pyramimonadaceae_tb_occur <- Pyramimonadaceae_tb[,1:91]
Pyramimonadaceae_tb_occur_len<-length(Pyramimonadaceae_tb_occur[,colSums(Pyramimonadaceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Pyramimonadaceae",samples_per_class=Pyramimonadaceae_tb_occur_len))

other_plastids_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "other_plastids"),]
other_plastids_tb_occur <- other_plastids_tb[,1:91]
other_plastids_tb_occur_len<-length(other_plastids_tb_occur[,colSums(other_plastids_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="other_plastids",samples_per_class=other_plastids_tb_occur_len))

Cyanobacteria_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Cyanobacteria"),]
Cyanobacteria_tb_occur <- Cyanobacteria_tb[,1:91]
Cyanobacteria_tb_occur_len<-length(Cyanobacteria_tb_occur[,colSums(Cyanobacteria_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Cyanobacteria",samples_per_class=Cyanobacteria_tb_occur_len))

occurrence_counts_phototrophs
#row.names(occurrence_counts_phototrophs)<-occurrence_counts_phototrophs$group
occurrence_counts_phototrophs<-as.data.frame(occurrence_counts_phototrophs)
```


**Absolute values**

```{r aggregate_non.raref_16S, echo=FALSE, results="hide", warning=FALSE}
class_summary_reads_per_class<-aggregate(rowSums(tb16_phototrophs[1:91]), list(tb16_phototrophs$class_A), sum)
# count the different groups

class_summary_otus_per_class<-aggregate(rowSums(tb16_phototrophs[1:91]), list(tb16_phototrophs$class_A), length)

## READS PER CLASS ##
attach(class_summary_reads_per_class)
class_summary_reads_per_class_order<-class_summary_reads_per_class[order(-x),]
detach(class_summary_reads_per_class)
class_summary_reads_per_class_order

#fix column names
row.names(class_summary_reads_per_class_order)<-class_summary_reads_per_class_order[,1]
class_summary_reads_per_class_order<-class_summary_reads_per_class_order[c(-1)]
colnames(class_summary_reads_per_class_order)<-c("reads_per_class")

## OTUs PER CLASS ##
attach(class_summary_otus_per_class)
class_summary_otus_per_class_order<-class_summary_otus_per_class[order(-x),]
detach(class_summary_otus_per_class)
class_summary_otus_per_class_order

row.names(class_summary_otus_per_class_order)<-class_summary_otus_per_class_order[,1]
class_summary_otus_per_class_order<-class_summary_otus_per_class_order[c(-1)]
colnames(class_summary_otus_per_class_order)<-c("OTUs_per_class")

#class_summary_reads<-aggregate(sum~class, data=otutab_full_wTax, FUN="sum") 
# sum reads different groups
```

```{r create_table_merging_#OTUs_#reads_#samples_non.raref_16S, echo=FALSE, warning=F}
tb16S_OTUs_reads_samples <- merge(class_summary_reads_per_class_order, class_summary_otus_per_class_order, by="row.names")

row.names(tb16S_OTUs_reads_samples)<-tb16S_OTUs_reads_samples[,1]

tb16S_OTUs_reads_samples<-tb16S_OTUs_reads_samples[,-1]
colnames(tb16S_OTUs_reads_samples)<-c("reads_per_class","OTUs_per_class")
#tb16S_OTUs_reads_samples[1:5,1:2]

tb16S_OTUs_reads_samples<-tb16S_OTUs_reads_samples[order(tb16S_OTUs_reads_samples$reads_per_class, tb16S_OTUs_reads_samples$OTUs_per_class, decreasing = T), c(1,2)]

#add no. of samples per class
tb16S_OTUs_reads_samples<-merge(tb16S_OTUs_reads_samples, occurrence_counts_phototrophs, by.x="row.names", by.y="group", all.x=TRUE)

row.names(tb16S_OTUs_reads_samples)<-tb16S_OTUs_reads_samples[,1]
tb16S_OTUs_reads_samples<-tb16S_OTUs_reads_samples[,-1]
tb16S_OTUs_reads_samples
```


\
**Relative values**

```{r OTUs_vs_reads_relative_values_18S_non.raref_16S, echo=FALSE, warning=F}

tb16S_OTUs_reads_samples_rel_abund <- tb16S_OTUs_reads_samples

tb16S_OTUs_reads_samples_rel_abund$reads_per_class<-(tb16S_OTUs_reads_samples_rel_abund$reads_per_class*100)/colSums(tb16S_OTUs_reads_samples)[1]
tb16S_OTUs_reads_samples_rel_abund$OTUs_per_class<-(tb16S_OTUs_reads_samples_rel_abund$OTUs_per_class*100)/colSums(tb16S_OTUs_reads_samples)[2]
tb16S_OTUs_reads_samples_rel_abund$samples_per_class<-(tb16S_OTUs_reads_samples_rel_abund$samples_per_class*100)/91

colSums(tb16S_OTUs_reads_samples_rel_abund)
tb16S_OTUs_reads_samples_rel_abund
```

\
\

Reads per class vs. OTUs per class:

```{r OTUs_vs_reads_18S_v2_non.raref_16S, echo=FALSE, warnings=F, warning=F}
ggplot(tb16S_OTUs_reads_samples_rel_abund) +
  geom_point(aes(OTUs_per_class, reads_per_class)) +
  geom_text_repel(aes(OTUs_per_class, reads_per_class, label = rownames(tb16S_OTUs_reads_samples)), size=4, force = 3, box.padding = unit(0.1, "lines"), point.padding = unit(0.02, 'lines')) + labs(title="[B7] MP 16S: reads per class vs. OTUs per class", x="\ndiversity (no. of OTUs in %)", y="abundance (no. of reads in %)\n") + scale_x_log10(limits = c(-1,200), breaks = c(0,0.1,1,10,100)) + scale_y_log10(limits = c(-1,200), breaks = c(0,0.1,1,10,100))  + theme(axis.text=element_text(size=14), axis.title=element_text(size=16))
```

\
\

Reads per class vs. samples in which they occurr:

```{r reads_vs_samples_18S_non.raref_16S, echo=FALSE, warning=F}
ggplot(tb16S_OTUs_reads_samples_rel_abund) +
  geom_point(aes(samples_per_class,reads_per_class)) +
  geom_text_repel(aes(samples_per_class,reads_per_class, label = rownames(tb16S_OTUs_reads_samples)), size=4, force = 8, box.padding = unit(0.1, "lines"), point.padding = unit(0.2, 'lines')) + scale_x_continuous(limits = c(0, 116), breaks=c(25,50,75,100)) + labs(title="[B8] MP 16S: reads per class vs. samples in which they occur", x="\nsamples in which they occurr (%)", y="abundance (no. of reads in %)\n") + scale_y_log10(limits = c(-1,200), breaks = c(0,0.1,1,10,100)) + theme(axis.text=element_text(size=14), axis.title=element_text(size=16))
```

\
\
\
\

```{r cyanobacteria_histogram_non.raref_16S, echo=F, results="hide", warning=F}
cyano_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Cyanobacteria"),]

class_summary_reads_per_class_cyano<-aggregate(rowSums(cyano_tb[1:91]), list(cyano_tb$class_B), sum)
# count the different groups
class_summary_otus_per_class_cyano<-aggregate(rowSums(cyano_tb[1:91]), list(cyano_tb$class_B), length)

## READS PER CLASS ##
attach(class_summary_reads_per_class_cyano)
class_summary_reads_per_class_cyano_order<-class_summary_reads_per_class_cyano[order(-x),]
detach(class_summary_reads_per_class_cyano)
class_summary_reads_per_class_cyano_order

#fix column names
row.names(class_summary_reads_per_class_cyano_order)<-class_summary_reads_per_class_cyano_order[,1]
class_summary_reads_per_class_cyano_order<-class_summary_reads_per_class_cyano_order[c(-1)]
colnames(class_summary_reads_per_class_cyano_order)<-c("reads_per_class")


## OTUs PER CLASS ##
attach(class_summary_otus_per_class_cyano)
class_summary_otus_per_class_cyano_order<-class_summary_otus_per_class_cyano[order(-x),]
detach(class_summary_otus_per_class_cyano)
class_summary_otus_per_class_cyano_order

row.names(class_summary_otus_per_class_cyano_order)<-class_summary_otus_per_class_cyano_order[,1]
class_summary_otus_per_class_cyano_order<-class_summary_otus_per_class_cyano_order[c(-1)]
colnames(class_summary_otus_per_class_cyano_order)<-c("OTUs_per_class")

# create_table_merging_#OTUs_#reads
cyano_OTUs_reads <- merge(class_summary_reads_per_class_cyano_order, class_summary_otus_per_class_cyano_order, by="row.names")

row.names(cyano_OTUs_reads)<-cyano_OTUs_reads[,1]
cyano_OTUs_reads<-cyano_OTUs_reads[,-1]
colnames(cyano_OTUs_reads)<-c("reads_per_class","OTUs_per_class")
cyano_OTUs_reads[1:3,1:2]

cyano_OTUs_reads<-cyano_OTUs_reads[order(cyano_OTUs_reads$reads_per_class, cyano_OTUs_reads$OTUs_per_class, decreasing = T), c(1,2)]
```

\
Absolute values of cyanobacteria groups richness and abundance:
```{r cyanos_abs_values_non.raref, echo=F}
cyano_OTUs_reads
```

```{r cyanos_rel_values_non.raref, echo=F, results="hide"}
#compute relative values
cyano_OTUs_reads_rel_abund <- cyano_OTUs_reads

cyano_OTUs_reads_rel_abund$reads_per_class<-(cyano_OTUs_reads_rel_abund$reads_per_class*100)/colSums(cyano_OTUs_reads)[1]
cyano_OTUs_reads_rel_abund$OTUs_per_class<-(cyano_OTUs_reads_rel_abund$OTUs_per_class*100)/colSums(cyano_OTUs_reads)[2]

colSums(cyano_OTUs_reads_rel_abund)
rownames(cyano_OTUs_reads_rel_abund) = c("Prochlorococcus", "Synechococcus", "Other cyanobacteria")
cyano_OTUs_reads_rel_abund

cyano_OTUs_reads_rel_abund["cyano_group"]<-NA
cyano_OTUs_reads_rel_abund$cyano_group<-c("Prochlorococcus", "Synechococcus", "Other cyanobacteria")

cyano_OTUs_reads_rel_abund2<-read.table("input_99/cyano_histograms_data_non.raref.txt", head=TRUE)
```


Relative values of Cyanobacteria groups richness and abundance:

```{r print_rel_values_non.raref, echo=F}
cyano_OTUs_reads_rel_abund[,1:2]

ggplot(cyano_OTUs_reads_rel_abund2, aes(x=group, y=value, fill=data)) + 
  geom_bar(position="dodge", stat="identity") +
  scale_y_continuous(limits=c(-1,100), breaks=c(25,50,75,100)) +
  scale_fill_manual(values=c("#000000", "cadetblue3")) +
  labs(x="", y="% \n") + 
  theme(axis.text=element_text(size=14), axis.title=element_text(size=16)) + 
  theme(legend.title = element_blank(), legend.position=c(0.85,0.85), legend.text=element_text(size=14)) +
  theme(axis.title.y = element_text(size = rel(1.5), angle = 0, vjust=0.97))
```



\
**PROTISTS**

```{r dim_merge2_16S.protists_non.raref, echo=F}
#selection of phototrophs:
tb16_phototrophs <- tb16_phototrophs[which(tb16_phototrophs$class_A != "Cyanobacteria"),]

nrow(tb16_phototrophs)
sum(colSums(tb16_phototrophs[,1:91]))

#dim(tb16_phototrophs)
#tb16_phototrophs[1:5,1:5]
```

```{r count_samples_per_group_16S.protists_non.raref, echo=F, warning=FALSE, results="hide"}
occurrence_counts_phototrophs<-data.table()

#create a table per group and count in how many samples they occur. 
Dinophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Dinophyceae"),]
Dinophyceae_tb_occur <- Dinophyceae_tb[,1:91]
Dinophyceae_tb_occur_len<-length(Dinophyceae_tb_occur[,colSums(Dinophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Dinophyceae",samples_per_class=Dinophyceae_tb_occur_len))

Prasinophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "other_Prasinophyceae"),]
Prasinophyceae_tb_occur <- Prasinophyceae_tb[,1:91]
Prasinophyceae_tb_occur_len<-length(Prasinophyceae_tb_occur[,colSums(Prasinophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="other_Prasinophyceae",samples_per_class=Prasinophyceae_tb_occur_len))

Chrysophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Chrysophyceae"),]
Chrysophyceae_tb_occur <- Chrysophyceae_tb[,1:91]
Chrysophyceae_tb_occur_len<-length(Chrysophyceae_tb_occur[,colSums(Chrysophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Chrysophyceae",samples_per_class=Chrysophyceae_tb_occur_len))

Pelagophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Pelagophyceae"),]
Pelagophyceae_tb_occur <- Pelagophyceae_tb[,1:91]
Pelagophyceae_tb_occur_len<-length(Pelagophyceae_tb_occur[,colSums(Pelagophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Pelagophyceae",samples_per_class=Pelagophyceae_tb_occur_len))

Dictyochophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Dictyochophyceae"),]
Dictyochophyceae_tb_occur <- Dictyochophyceae_tb[,1:91]
Dictyochophyceae_tb_occur_len<-length(Dictyochophyceae_tb_occur[,colSums(Dictyochophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Dictyochophyceae",samples_per_class=Dictyochophyceae_tb_occur_len))

Cryptophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Cryptophyceae"),]
Cryptophyceae_tb_occur <- Cryptophyceae_tb[,1:91]
Cryptophyceae_tb_occur_len<-length(Cryptophyceae_tb_occur[,colSums(Cryptophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Cryptophyceae",samples_per_class=Cryptophyceae_tb_occur_len))

Bacillariophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Bacillariophyceae"),]
Bacillariophyceae_tb_occur <- Bacillariophyceae_tb[,1:91]
Bacillariophyceae_tb_occur_len<-length(Bacillariophyceae_tb_occur[,colSums(Bacillariophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Bacillariophyceae",samples_per_class=Bacillariophyceae_tb_occur_len))

Chlorarachniophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Chlorarachniophyceae"),]
Chlorarachniophyceae_tb_occur <- Chlorarachniophyceae_tb[,1:91]
Chlorarachniophyceae_tb_occur_len<-length(Chlorarachniophyceae_tb_occur[,colSums(Chlorarachniophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Chlorarachniophyceae",samples_per_class=Chlorarachniophyceae_tb_occur_len))

Bolidophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Bolidophyceae"),]
Bolidophyceae_tb_occur <- Bolidophyceae_tb[,1:91]
Bolidophyceae_tb_occur_len<-length(Bolidophyceae_tb_occur[,colSums(Bolidophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Bolidophyceae",samples_per_class=Bolidophyceae_tb_occur_len))

Pinguiophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Pinguiophyceae"),]
Pinguiophyceae_tb_occur <- Pinguiophyceae_tb[,1:91]
Pinguiophyceae_tb_occur_len<-length(Pinguiophyceae_tb_occur[,colSums(Pinguiophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Pinguiophyceae",samples_per_class=Pinguiophyceae_tb_occur_len))

Prymnesiophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Prymnesiophyceae"),]
Prymnesiophyceae_tb_occur <- Prymnesiophyceae_tb[,1:91]
Prymnesiophyceae_tb_occur_len<-length(Prymnesiophyceae_tb_occur[,colSums(Prymnesiophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Prymnesiophyceae",samples_per_class=Prymnesiophyceae_tb_occur_len))

Mamiellophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Mamiellophyceae"),]
Mamiellophyceae_tb_occur <- Mamiellophyceae_tb[,1:91]
Mamiellophyceae_tb_occur_len<-length(Mamiellophyceae_tb_occur[,colSums(Mamiellophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Mamiellophyceae",samples_per_class=Mamiellophyceae_tb_occur_len))

Eustigmatophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Eustigmatophyceae"),]
Eustigmatophyceae_tb_occur <- Eustigmatophyceae_tb[,1:91]
Eustigmatophyceae_tb_occur_len<-length(Eustigmatophyceae_tb_occur[,colSums(Eustigmatophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Eustigmatophyceae",samples_per_class=Eustigmatophyceae_tb_occur_len))

Chlorophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Chlorophyceae"),]
Chlorophyceae_tb_occur <- Chlorophyceae_tb[,1:91]
Chlorophyceae_tb_occur_len<-length(Chlorophyceae_tb_occur[,colSums(Chlorophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Chlorophyceae",samples_per_class=Chlorophyceae_tb_occur_len))

Ulvophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Ulvophyceae"),]
Ulvophyceae_tb_occur <- Ulvophyceae_tb[,1:91]
Ulvophyceae_tb_occur_len<-length(Ulvophyceae_tb_occur[,colSums(Ulvophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Ulvophyceae",samples_per_class=Ulvophyceae_tb_occur_len))

Raphydophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Raphydophyceae"),]
Raphydophyceae_tb_occur <- Raphydophyceae_tb[,1:91]
Raphydophyceae_tb_occur_len<-length(Raphydophyceae_tb_occur[,colSums(Raphydophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Raphydophyceae",samples_per_class=Raphydophyceae_tb_occur_len))

Trebouxiophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Trebouxiophyceae"),]
Trebouxiophyceae_tb_occur <- Trebouxiophyceae_tb[,1:91]
Trebouxiophyceae_tb_occur_len<-length(Trebouxiophyceae_tb_occur[,colSums(Trebouxiophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Trebouxiophyceae",samples_per_class=Trebouxiophyceae_tb_occur_len))

Phaeophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Phaeophyceae"),]
Phaeophyceae_tb_occur <- Phaeophyceae_tb[,1:91]
Phaeophyceae_tb_occur_len<-length(Phaeophyceae_tb_occur[,colSums(Phaeophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Phaeophyceae",samples_per_class=Phaeophyceae_tb_occur_len))

Phaeothamniophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Phaeothamniophyceae"),]
Phaeothamniophyceae_tb_occur <- Phaeothamniophyceae_tb[,1:91]
Phaeothamniophyceae_tb_occur_len<-length(Phaeothamniophyceae_tb_occur[,colSums(Phaeothamniophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Phaeothamniophyceae",samples_per_class=Phaeothamniophyceae_tb_occur_len))

Xanthophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Xanthophyceae"),]
Xanthophyceae_tb_occur <- Xanthophyceae_tb[,1:91]
Xanthophyceae_tb_occur_len<-length(Xanthophyceae_tb_occur[,colSums(Xanthophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Xanthophyceae",samples_per_class=Xanthophyceae_tb_occur_len))

Chlorodendrophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Chlorodendrophyceae"),]
Chlorodendrophyceae_tb_occur <- Chlorodendrophyceae_tb[,1:91]
Chlorodendrophyceae_tb_occur_len<-length(Chlorodendrophyceae_tb_occur[,colSums(Chlorodendrophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Chlorodendrophyceae",samples_per_class=Chlorodendrophyceae_tb_occur_len))

IncertaeSedis_Archaeplastida_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "IncertaeSedis_Archaeplastida"),]
IncertaeSedis_Archaeplastida_tb_occur <- IncertaeSedis_Archaeplastida_tb[,1:91]
IncertaeSedis_Archaeplastida_tb_occur_len<-length(IncertaeSedis_Archaeplastida_tb_occur[,colSums(IncertaeSedis_Archaeplastida_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="IncertaeSedis_Archaeplastida",samples_per_class=IncertaeSedis_Archaeplastida_tb_occur_len))

Nephroselmidophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Nephroselmidophyceae"),]
Nephroselmidophyceae_tb_occur <- Nephroselmidophyceae_tb[,1:91]
Nephroselmidophyceae_tb_occur_len<-length(Nephroselmidophyceae_tb_occur[,colSums(Nephroselmidophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Nephroselmidophyceae",samples_per_class=Nephroselmidophyceae_tb_occur_len))

Pavlovophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Pavlovophyceae"),]
Pavlovophyceae_tb_occur <- Pavlovophyceae_tb[,1:91]
Pavlovophyceae_tb_occur_len<-length(Pavlovophyceae_tb_occur[,colSums(Pavlovophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Pavlovophyceae",samples_per_class=Pavlovophyceae_tb_occur_len))

Rhodophyceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Rhodophyceae"),]
Rhodophyceae_tb_occur <- Rhodophyceae_tb[,1:91]
Rhodophyceae_tb_occur_len<-length(Rhodophyceae_tb_occur[,colSums(Rhodophyceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Rhodophyceae",samples_per_class=Rhodophyceae_tb_occur_len))

Rappemonads_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Rappemonads"),]
Rappemonads_tb_occur <- Rappemonads_tb[,1:91]
Rappemonads_tb_occur_len<-length(Rappemonads_tb_occur[,colSums(Rappemonads_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Rappemonads",samples_per_class=Rappemonads_tb_occur_len))

MOCH_1_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "MOCH_1"),]
MOCH_1_tb_occur <- MOCH_1_tb[,1:91]
MOCH_1_tb_occur_len<-length(MOCH_1_tb_occur[,colSums(MOCH_1_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="MOCH_1",samples_per_class=MOCH_1_tb_occur_len))

MOCH_2_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "MOCH_2"),]
MOCH_2_tb_occur <- MOCH_2_tb[,1:91]
MOCH_2_tb_occur_len<-length(MOCH_2_tb_occur[,colSums(MOCH_2_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="MOCH_2",samples_per_class=MOCH_2_tb_occur_len))

MOCH_5_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "MOCH_5"),]
MOCH_5_tb_occur <- MOCH_5_tb[,1:91]
MOCH_5_tb_occur_len<-length(MOCH_5_tb_occur[,colSums(MOCH_5_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="MOCH_5",samples_per_class=MOCH_5_tb_occur_len))

Prasinophyceae_clade_VII_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Prasinophyceae_clade-VII"),]
Prasinophyceae_clade_VII_tb_occur <- Prasinophyceae_clade_VII_tb[,1:91]
Prasinophyceae_clade_VII_tb_occur_len<-length(Prasinophyceae_clade_VII_tb_occur[,colSums(Prasinophyceae_clade_VII_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Prasinophyceae_clade-VII",samples_per_class=Prasinophyceae_clade_VII_tb_occur_len))

Prasinophyceae_clade_IX_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Prasinophyceae_clade-IX"),]
Prasinophyceae_clade_IX_tb_occur <- Prasinophyceae_clade_IX_tb[,1:91]
Prasinophyceae_clade_IX_tb_occur_len<-length(Prasinophyceae_clade_IX_tb_occur[,colSums(Prasinophyceae_clade_IX_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Prasinophyceae_clade-IX",samples_per_class=Prasinophyceae_clade_IX_tb_occur_len))

Pyramimonadaceae_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "Pyramimonadaceae"),]
Pyramimonadaceae_tb_occur <- Pyramimonadaceae_tb[,1:91]
Pyramimonadaceae_tb_occur_len<-length(Pyramimonadaceae_tb_occur[,colSums(Pyramimonadaceae_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="Pyramimonadaceae",samples_per_class=Pyramimonadaceae_tb_occur_len))

other_plastids_tb <- tb16_phototrophs[which(tb16_phototrophs$class_A == "other_plastids"),]
other_plastids_tb_occur <- other_plastids_tb[,1:91]
other_plastids_tb_occur_len<-length(other_plastids_tb_occur[,colSums(other_plastids_tb_occur) > 0])
occurrence_counts_phototrophs<-rbind(occurrence_counts_phototrophs,data.table(group="other_plastids",samples_per_class=other_plastids_tb_occur_len))

occurrence_counts_phototrophs
#row.names(occurrence_counts_phototrophs)<-occurrence_counts_phototrophs$group
occurrence_counts_phototrophs<-as.data.frame(occurrence_counts_phototrophs)
```


**Absolute values**

```{r aggregate_16S_protists_non.raref, echo=FALSE, results="hide", warning=FALSE}
class_summary_reads_per_class<-aggregate(rowSums(tb16_phototrophs[1:91]), list(tb16_phototrophs$class_A), sum)
# count the different groups

class_summary_otus_per_class<-aggregate(rowSums(tb16_phototrophs[1:91]), list(tb16_phototrophs$class_A), length)

## READS PER CLASS ##
attach(class_summary_reads_per_class)
class_summary_reads_per_class_order<-class_summary_reads_per_class[order(-x),]
detach(class_summary_reads_per_class)
class_summary_reads_per_class_order

#fix column names
row.names(class_summary_reads_per_class_order)<-class_summary_reads_per_class_order[,1]
class_summary_reads_per_class_order<-class_summary_reads_per_class_order[c(-1)]
colnames(class_summary_reads_per_class_order)<-c("reads_per_class")

## OTUs PER CLASS ##
attach(class_summary_otus_per_class)
class_summary_otus_per_class_order<-class_summary_otus_per_class[order(-x),]
detach(class_summary_otus_per_class)
class_summary_otus_per_class_order

row.names(class_summary_otus_per_class_order)<-class_summary_otus_per_class_order[,1]
class_summary_otus_per_class_order<-class_summary_otus_per_class_order[c(-1)]
colnames(class_summary_otus_per_class_order)<-c("OTUs_per_class")

#class_summary_reads<-aggregate(sum~class, data=otutab_full_wTax, FUN="sum") 
# sum reads different groups
```

```{r create_table_merging_#OTUs_#reads_#samples_16S_protists_non.raref, echo=FALSE, warning=F}
tb16S_OTUs_reads_samples <- merge(class_summary_reads_per_class_order, class_summary_otus_per_class_order, by="row.names")

row.names(tb16S_OTUs_reads_samples)<-tb16S_OTUs_reads_samples[,1]

tb16S_OTUs_reads_samples<-tb16S_OTUs_reads_samples[,-1]
colnames(tb16S_OTUs_reads_samples)<-c("reads_per_class","OTUs_per_class")
#tb16S_OTUs_reads_samples[1:5,1:2]

tb16S_OTUs_reads_samples<-tb16S_OTUs_reads_samples[order(tb16S_OTUs_reads_samples$reads_per_class, tb16S_OTUs_reads_samples$OTUs_per_class, decreasing = T), c(1,2)]

#add no. of samples per class
tb16S_OTUs_reads_samples<-merge(tb16S_OTUs_reads_samples, occurrence_counts_phototrophs, by.x="row.names", by.y="group", all.x=TRUE)

row.names(tb16S_OTUs_reads_samples)<-tb16S_OTUs_reads_samples[,1]
tb16S_OTUs_reads_samples<-tb16S_OTUs_reads_samples[,-1]
tb16S_OTUs_reads_samples
```


\
**Relative values**

```{r OTUs_vs_reads_relative_values_16S_non.raref, echo=FALSE, warning=F}

tb16S_OTUs_reads_samples_rel_abund <- tb16S_OTUs_reads_samples

tb16S_OTUs_reads_samples_rel_abund$reads_per_class<-(tb16S_OTUs_reads_samples_rel_abund$reads_per_class*100)/colSums(tb16S_OTUs_reads_samples)[1]
tb16S_OTUs_reads_samples_rel_abund$OTUs_per_class<-(tb16S_OTUs_reads_samples_rel_abund$OTUs_per_class*100)/colSums(tb16S_OTUs_reads_samples)[2]
tb16S_OTUs_reads_samples_rel_abund$samples_per_class<-(tb16S_OTUs_reads_samples_rel_abund$samples_per_class*100)/91

colSums(tb16S_OTUs_reads_samples_rel_abund)
tb16S_OTUs_reads_samples_rel_abund
```

\
\

Reads per class vs. OTUs per class:

```{r OTUs_vs_reads_16S_v2_non.raref, echo=FALSE, warnings=F, warning=F}
ggplot(tb16S_OTUs_reads_samples_rel_abund) +
  geom_point(aes(OTUs_per_class, reads_per_class)) +
  geom_text_repel(aes(OTUs_per_class, reads_per_class, label = rownames(tb16S_OTUs_reads_samples)), size=4, force = 3, box.padding = unit(0.1, "lines"), point.padding = unit(0.02, 'lines')) + labs(title="[B7.2] MP 16S: reads per class vs. OTUs per class", x="\ndiversity (no. of OTUs in %)", y="abundance (no. of reads in %)\n") + scale_x_log10(limits = c(-1,200), breaks = c(0,0.1,1,10,100)) + scale_y_log10(limits = c(-1,200), breaks = c(0,0.1,1,10,100))  + theme(axis.text=element_text(size=14), axis.title=element_text(size=16))
```

\
\

Reads per class vs. samples in which they occurr:

```{r reads_vs_samples_16S_non.raref, echo=FALSE, warning=F}
ggplot(tb16S_OTUs_reads_samples_rel_abund) +
  geom_point(aes(samples_per_class,reads_per_class)) +
  geom_text_repel(aes(samples_per_class,reads_per_class, label = rownames(tb16S_OTUs_reads_samples)), size=4, force = 8, box.padding = unit(0.1, "lines"), point.padding = unit(0.2, 'lines')) + scale_x_continuous(limits = c(0, 116), breaks=c(25,50,75,100)) + labs(title="[B8.2] MP 16S: reads per class vs. samples in which they occur", x="\nsamples in which they occurr (%)", y="abundance (no. of reads in %)\n") + scale_y_log10(limits = c(-1,200), breaks = c(0,0.1,1,10,100)) + theme(axis.text=element_text(size=14), axis.title=element_text(size=16))
```
\








